### 2014-05-17 paper.tex
paper skeleton + NLP demo
Fixes #1. I didn't try pdflatex on this yet so it may not build.

\documentclass[a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{authblk}
\usepackage{dblfloatfix}
%\usepackage{hyperref}
\usepackage{minted}
\usepackage[numbers,square]{natbib}
\usepackage{xcolor}
\usepackage{xspace}
\usepackage{url}

\definecolor{rulecolor}{rgb}{0.80,0.80,0.80}
\newminted{python}{frame=single,rulecolor=\color{rulecolor}}

\title{scikit-learn is the greatest}

\date{}

\begin{document}

\maketitle

% Lars: the following text presumes that NumPy, estimators, and the
% X/y convention for supervised learning have already been introduced.
\section{Text and language processing}

Scikit-learn was designed around NumPy principles and data structures,
so virtually all estimators take arrays of numbers as input
and produce arrays as output.
That doesn't mean that it is limited to traditional number crunching domains,
though.
The library has been successfully employed in text classification
and natural language processing problems,
and utility code is provided to translate the unstructured, symbolic data
encountered in these domains
to the rigid matrix/vector structure that estimators expect.

In the NLP application domain,
features are typically 
where features are typically frequencies or boolean indicators
of very large numbers of rare events, such as
``document contains the term \textsf{v1agr@}'' (spam filtering)
or ``next token matches the Roman numeral regex'' (named-entity recognition).
To ensure scalability in the face of such data,
SciPy's sparse matrix data structures are supported by many estimators.\footnote{
  Support for these is added on a by-need basis,
  because using sparse data structures puts constraints on algorithm design
  and implementation.
}

The following example script is a fully functioning
sentiment polarity classifier for movie reviews:
given a review, it outputs the probability that the review is positive
about the movie it concerns.
The script demonstrates the use of pipelines,
and also shows how machine learning ties into Python's
``batteries included'' philosophy:
downloading, unpacking and learning from a dataset
are all done in one programming language.

% XXX make_pipeline is not in 0.14.1, we should release 0.15 sometime
% The example can be made a bit shorter by skipping the movie_reviews dir;
% the tarball from Pang et al. contain top-level cruft.
\begin{pythoncode}
import os.path
import tarfile
from urllib import urlretrieve
from sklearn.datasets import load\_files
from sklearn.feature\_extraction.text import TfidfVectorizer
from sklearn.linear\_models import LogisticRegression
from sklearn.pipeline import make\_pipeline

temp, \_ = urlretrieve('http://www.cs.cornell.edu/people/pabo/movie-review-data/review\_polarity.tar.gz')
with tarfile.open(temp) as tar:
    tar.extractall(path='movie\_reviews')
data = load\_files(os.path.join('movie\_reviews', 'txt\_sentoken'))

clf = make\_pipeline(TfidfVectorizer(min\_df=2, dtype=float,
                                    sublinear\_tf=True, ngram\_range=(1, 2),
                                    strip\_accents='unicode'),
                     LogisticRegression(random\_state=623, C=5000))
clf.fit(data.data, data.target)
\end{pythoncode}

This script first fetches the hand-labeled movie review dataset from
\citet{pang2004} and unpacks it to disk using standard Python libraries.
It then loads these into memory using the scikit-learn function
\texttt{load\_files}, makes a pipeline of a \textsf{tf--idf} feature extractor
\citep{rennie2003tackling} and logistic regression,
and finally trains the model.
The parameters for the feature extractor and the LogReg classifier
were tuned offline; the random-number generator seed is set by hand
to get reproducible results.

Does this classifier work?
We can find out by classifying some of our own movie reviews.\footnote{
  The code can be tried interactively by simply copy-pasting it
  into an interactive Python interpreter or IPython \citep{perez2007ipython}.
}

\begin{pythoncode}
>>> clf.predict\_proba(["This movie is the worst I ever saw.",
...                    "Shawshank Redemption, eat your heart out!"])
array([[  9.99293000e-01,   7.07000280e-04],
       [  1.68757754e-01,   8.31242246e-01]])
\end{pythoncode}

Here, we asked the model to predict the probability of a review being positive
for a batch of two reviews. It returns the negative probability in the first
column, the positive probability in the second.
As we can see, the first review gets a 99.9\% probability for being negative,
while the second has an 83\% probability of being positive.

Of course, that's just anecdotal evidence. We can get a figure for the
classifier's accuracy by retraining it in a five-fold cross-validation scheme:

\begin{pythoncode}
>>> from sklearn.cross\_validation import cross\_val\_score
>>> score = cross\_val\_score(clf, data.data, data.target, cv=5)
>>> score
array([ 0.91  ,  0.8825,  0.88  ,  0.8775,  0.86  ])
>>> score.mean()
0.88200000000000001
\end{pythoncode}

\end{document}

### 2014-05-18 article.tex
Iteration on the paper, and an example on wage

\documentclass{article}
\usepackage{magazine-style}

\begin{document}

\title{Scikit-learn: machine learning without learning the machinery}

\author{The authors}

\maketitle

\section{Introduction}

\subsection{Project vision}

Bridging the gap between machine learning research and applications

\cite{pedregosa2011}

Impact on education (documentation)

Striving for ease-of-use and quality

Importance of API design \cite{buitinck2013ecml}

\subsection{The python data ecosystem}

In a data analysis pipeline, machine learning is only a tiny bit of the
pipeline, although it is critical.

Great scientific and numeric ecosystem
\cite{oliphant2007python,varoquaux2013scipy}, but also text processing,
webservers

Numpy arrays \cite{vanderwalt2011}, Scipy, matplotlib. Cython
\cite{behnel2011cython}

Pandas for columnar data, scikit-image, NLTK.

Ipython for interactive work \cite{perez2007ipython}

\subsection{Some history of the project}

PhD project of David Cournapeau and Matthieu Brucher, Parietal-INRIA
comes along,... 2010 sprint,

LibSVM a good start
\cite{chang2011libsvm}

\section{Learning with scikit-learn}

\subsection{The data matrix}

Blahblah

\subsection{Supervised models: learning to predict}

Blahblah

\subsection{Model evaluation and parameter selection}

Model evaluation:
cross-validation: the concept, cross-val iterators, and cross\_val\_score

Problem of multiple metrics.

CV objects: GridSearchCV and RandomSearchCV, as well as FooBarCV objects.

\subsection{Unsupervised models: learning to transform}

Unsupervised learning: a variety of usage pattern.

Many things can be seen as Transformers

Introduce the pipeline

\section{In practice}

\subsection{A simple text-mining example}

Example from Lars here

\subsection{What about big data?}

Partial fit, random projections, HashingVectorizer

\section{Nurturing an open source project}

Goal: enable anybody to contribute, have a controlled process, grow.

Github, tests (link to travis)

Difficulty of getting credit, of rewarding properly the long tail of small
contributors, of finding funding. Problem of brain drain.

Difficulty of project scope (ideally: first cover all of statistical
learning classics \cite{elemstatlearn}) and default parameters

We have found that implementing, even a standard algorithm, really well,
can require a lot of domain knowledge. Thus it is natural that specific
libraries span up to solve it. Some adopt scikit-learn API and standard,
and we hope that scikit-learn has a structural effect on the environment.

\small
\bibliography{paper}
\bibliographystyle{plain}

\end{document}

### 2014-05-18 wage_data.py

@@ -1,0 +1,131 @@
+import urllib, os # Standard library imports
+
+import numpy as np
+import matplotlib.pyplot as plt
+
+from sklearn import svm, ensemble
+
+###############################################################################
+# Load the data
+import pandas
+
+if not os.path.exists('wages.txt'):
+    urllib.urlretrieve('http://lib.stat.cmu.edu/datasets/CPS_85_Wages',
+                       'wages.txt')
+
+names = [
+    'EDUCATION: Number of years of education',
+    'SOUTH: 1=Person lives in South, 0=Person lives elsewhere',
+    'SEX: 1=Female, 0=Male',
+    'EXPERIENCE: Number of years of work experience',
+    'UNION: (1=Union member, 0=Not union member)',
+    'WAGE: Wage (dollars per hour)',
+    'AGE: years',
+    'RACE: 1=Other, 2=Hispanic, 3=White',
+    'OCCUPATION: 1=Management, 2=Sales, 3=Clerical, 4=Service, 5=Professional, 6=Other',
+    'SECTOR: 0=Other, 1=Manufacturing, 2=Construction',
+    'MARR: 0=Unmarried,  1=Married',
+]
+
+short_names = [n.split(':')[0] for n in names]
+
+data = pandas.read_csv('wages.txt', skiprows=27, skipfooter=6, sep=None,
+       header=None)
+data.columns = short_names
+
+# In the color plots, crop at 30 to have more dynamical range on the
+# color
+wage_max = 30
+
+for name, learner in [('', None),
+                      (' linear SVM', svm.SVR(kernel='linear')),
+                      (' random forest', ensemble.RandomForestRegressor())]:
+    #-----------------------------------------------------------------------
+    # Plot a simple pair-wise plot
+    plt.figure(figsize=(6, 5))
+
+    ax_2d = plt.axes([.38, .33, .6, .6])
+
+    # Add random offset to limit overlap in points
+    n_sample = len(data)
+    plt.scatter(data['EXPERIENCE'] -.25 + .5 * np.random.random(n_sample),
+                data['EDUCATION'] -.25 + .5 * np.random.random(n_sample),
+                c=data['WAGE'], cmap=plt.cm.Blues, vmax=wage_max)
+
+    plt.xlabel('Years of work experience')
+    plt.ylabel('Years of education')
+    plt.yticks(size=10)
[…]

### 2014-05-18 article.tex
More detailed layout
merge paper.tex, article.tex

@@
-\documentclass{article}
+\documentclass[a4paper]{article}
+\usepackage[utf8]{inputenc}
+\usepackage{amsmath}
+\usepackage{amsfonts}
+\usepackage{authblk}
+\usepackage{dblfloatfix}
+%\usepackage{hyperref}
+\usepackage{minted}
+\usepackage[numbers,square]{natbib}
+\usepackage{xcolor}
+\usepackage{xspace}
+\usepackage{url}
 \usepackage{magazine-style}
 
-\begin{document}
+\definecolor{rulecolor}{rgb}{0.80,0.80,0.80}
+\newminted{python}{frame=single,rulecolor=\color{rulecolor}}
 
@@
 
+\begin{document}
+
 \maketitle
 
-\section{Introduction}
+\section{Scikit-learn: software to bridge a gap}
 
-\subsection{Project vision}
-
+\paragraph{Project vision}
+%
 Bridging the gap between machine learning research and applications
@@
 
-\subsection{The python data ecosystem}
-
+\paragraph{The python data ecosystem}
+%
 In a data analysis pipeline, machine learning is only a tiny bit of the
@@
 
-Numpy arrays \cite{vanderwalt2011}, Scipy, matplotlib. Cython
+NumPy arrays \cite{vanderwalt2011}, SciPy, matplotlib. Cython
 \cite{behnel2011cython}
@@
 
-Ipython for interactive work \cite{perez2007ipython}
+IPython for interactive work \cite{perez2007ipython}
 
-\subsection{Some history of the project}
-
+\paragraph{Some history of the project}
+%
 PhD project of David Cournapeau and Matthieu Brucher, Parietal-INRIA
@@
 
+\section{A brief introduction to machine learning}
+
+Discuss what we can learn from the wages examples. Introduce concepts of
+overfitting, model complexity
+
 \section{Learning with scikit-learn}
 
-\subsection{The data matrix}
-
+\paragraph{The data matrix}
+%
 Blahblah
 
-\subsection{Supervised models: learning to predict}
-
+\paragraph{Supervised models: learning to predict}
+%
 Blahblah
 
-\subsection{Model evaluation and parameter selection}
-
+\paragraph{Model evaluation and parameter selection}
+%
 Model evaluation:
@@
 
-\subsection{Unsupervised models: learning to transform}
-
+\paragraph{Unsupervised models: learning to transform}
+%
 Unsupervised learning: a variety of usage pattern.
@@
 
-\subsection{A simple text-mining example}
+\paragraph{A simple text-mining example}
+%
+% Lars: the following text presumes that NumPy, estimators, and the
+% X/y convention for supervised learning have already been introduced.
 
-Example from Lars here
+Scikit-learn was designed around NumPy principles and data structures,
+so virtually all estimators take arrays of numbers as input
+and produce arrays as output.
+That doesn't mean that it is limited to traditional number crunching domains,
+though.
+The library has been successfully employed in text classification
+and natural language processing problems,
+and utility code is provided to translate the unstructured, symbolic data
+encountered in these domains
+to the rigid matrix/vector structure that estimators expect.
 
-\subsection{What about big data?}
+In the NLP application domain,
+features are typically 
+where features are typically frequencies or boolean indicators
+of very large numbers of rare events, such as
+``document contains the term \textsf{v1agr@}'' (spam filtering)
+or ``next token matches the Roman numeral regex'' (named-entity recognition).
+To ensure scalability in the face of such data,
+SciPy's sparse matrix data structures are supported by many estimators.\footnote{
+  Support for these is added on a by-need basis,
+  because using sparse data structures puts constraints on algorithm design
+  and implementation.
+}
 
+The following example script is a fully functioning
+sentiment polarity classifier for movie reviews:
+given a review, it outputs the probability that the review is positive
+about the movie it concerns.
+The script demonstrates the use of pipelines,
+and also shows how machine learning ties into Python's
+``batteries included'' philosophy:
+downloading, unpacking and learning from a dataset
+are all done in one programming language.
+
+% XXX make_pipeline is not in 0.14.1, we should release 0.15 sometime
+% The example can be made a bit shorter by skipping the movie_reviews dir;
+% the tarball from Pang et al. contain top-level cruft.
+\begin{pythoncode}
+import os.path
+import tarfile
+from urllib import urlretrieve
+from sklearn.datasets import load\_files
+from sklearn.feature\_extraction.text import TfidfVectorizer
+from sklearn.linear\_models import LogisticRegression
+from sklearn.pipeline import make\_pipeline
+
+temp, \_ = urlretrieve('http://www.cs.cornell.edu/people/pabo/movie-review-data/review\_polarity.tar.gz')
+with tarfile.open(temp) as tar:
+    tar.extractall(path='movie\_reviews')
+data = load\_files(os.path.join('movie\_reviews', 'txt\_sentoken'))
+
+clf = make\_pipeline(TfidfVectorizer(min\_df=2, dtype=float,
+                                    sublinear\_tf=True, ngram\_range=(1, 2),
+                                    strip\_accents='unicode'),
+                     LogisticRegression(random\_state=623, C=5000))
+clf.fit(data.data, data.target)
+\end{pythoncode}
+
+This script first fetches the hand-labeled movie review dataset from
+\cite{pang2004} and unpacks it to disk using standard Python libraries.
+It then loads these into memory using the scikit-learn function
+\texttt{load\_files}, makes a pipeline of a \textsf{tf--idf} feature extractor
+\cite{rennie2003tackling} and logistic regression,
+and finally trains the model.
+The parameters for the feature extractor and the LogReg classifier
+were tuned offline; the random-number generator seed is set by hand
+to get reproducible results.
+
+Does this classifier work?
+We can find out by classifying some of our own movie reviews.\footnote{
+  The code can be tried interactively by simply copy-pasting it
+  into an interactive Python interpreter or IPython \cite{perez2007ipython}.
+}
+
+\begin{pythoncode}
+>>> clf.predict\_proba(["This movie is the worst I ever saw.",
+...                    "Shawshank Redemption, eat your heart out!"])
+array([[  9.99293000e-01,   7.07000280e-04],
+       [  1.68757754e-01,   8.31242246e-01]])
+\end{pythoncode}
+
+Here, we asked the model to predict the probability of a review being positive
+for a batch of two reviews. It returns the negative probability in the first
+column, the positive probability in the second.
+As we can see, the first review gets a 99.9\% probability for being negative,
+while the second has an 83\% probability of being positive.
+
+Of course, that's just anecdotal evidence. We can get a figure for the
+classifier's accuracy by retraining it in a five-fold cross-validation scheme:
+
+\begin{pythoncode}
+>>> from sklearn.cross\_validation import cross\_val\_score
+>>> score = cross\_val\_score(clf, data.data, data.target, cv=5)
+>>> score
+array([ 0.91  ,  0.8825,  0.88  ,  0.8775,  0.86  ])
+>>> score.mean()
+0.88200000000000001
+\end{pythoncode}
+
+\paragraph{What about big data?}
+%
 Partial fit, random projections, HashingVectorizer

### 2014-05-18 article.tex
Iter on paper
The abstract

@@
 
-\section{Scikit-learn: software to bridge a gap}
+\begin{abstract}
+Machine learning is a pervasive development at the intersection of
+statistics and computer science. While it could benefit to many
+data-related applications, the technicality of the research literature
+and the corresponding algorithms slows down its adoption. Scikit-learn is
+a open-source software project that aims to make machine learning
+accessible to all, whether it be in academia or in the industry. It
+fosters the general-purpose Python language, that enjoys a huge growth in
+the scientific world, and well as a striving ecosystem of contributors.
+Here we give a quick introduction to the scikit-learn as well as to
+machine-learning basics.
+\end{abstract}
+
+\section{A software project across communities}
 

### 2014-05-18 article.tex
Wording

@@
 Machine learning is a pervasive development at the intersection of
-statistics and computer science. While it could benefit to many
+statistics and computer science. While it can benefit many
 data-related applications, the technicality of the research literature

### 2014-05-18 article.tex
Iter on the paper

@@
 %\usepackage{hyperref}
+\usepackage{graphicx}
 \usepackage{minted}
@@
 %
+Scikit-learn was born from the observation that most standard
+machine-learning algorithms were out of reach of the users that could
+most benefit from them, such as researchers --biologists, climate
+scientists, experimental physicists--, developers in the industry --web
+services, domain-specific applications-- or any analyst without a strong
+training in statistics.
+%
+Implementations mostly consisted of scattered piece of code to download
+on researchers' web pages, with the notable exception a few
+statistic-specific environment such as the R language \cite{Rmanual}, or
+the Weka java library \cite{hall2009weka}.
+
 Bridging the gap between machine learning research and applications
@@
 
-\paragraph{The python data ecosystem}
+\paragraph{The Python data ecosystem}
 %
@@
 overfitting, model complexity
+
+\begin{figure*}
+    \includegraphics[width=.355\linewidth]{wage_data}%
+    \hspace*{-.025\linewidth}%
+    \hfill%
+    \includegraphics[width=.355\linewidth]{wage_data_random_forest}%
+    \hspace*{-.025\linewidth}%
+    \hfill%
+    \includegraphics[width=.355\linewidth]{wage_data_linear_svm}%
+
+\end{figure*}
 

### 2014-05-18 wage_data.py

@@ -44,7 +44,7 @@ for name, learner in [('', None),
     # Plot a simple pair-wise plot
     plt.figure(figsize=(6, 5))
 
-    ax_2d = plt.axes([.38, .33, .6, .6])
+    ax_2d = plt.axes([.35, .33, .6, .6])
 
     # Add random offset to limit overlap in points
     n_sample = len(data)
@@ -61,13 +61,17 @@ for name, learner in [('', None),
     ex_min, ex_max = plt.xlim()
     ed_min, ed_max = plt.ylim()
     cb = plt.colorbar()
+    # Squeeze the colorbar a bit left
+    bb = cb.ax.get_position()
+    bb.x0 -= .03
+    cb.ax.set_position(bb)
     cb.ax.set_ylabel('Wage')
     for l in cb.ax.get_yticklabels():
         l.update(dict(size=10))
 
     # Side plot as a function of education
-    ax_ed = plt.axes([.14, .33, .2, .6])
-    plt.plot(data['WAGE'], data['EDUCATION'], '+', label='Observed data')
+    ax_ed = plt.axes([.11, .33, .2, .6])
+    plt.plot(data['WAGE'], data['EDUCATION'], '+', label='Observed\n data')
     plt.ylim(ed_min, ed_max)
     plt.xlim(xmax=48)
     plt.ylabel('Years of education')
@@ -76,8 +80,8 @@ for name, learner in [('', None),
     plt.yticks(size=10)
 
     # Side plot as a function of experience
-    ax_ex = plt.axes([.38, .09, .48, .2])
-    plt.plot(data['EXPERIENCE'], data['WAGE'], '+', label='Observed data')
+    ax_ex = plt.axes([.35, .09, .48, .2])
+    plt.plot(data['EXPERIENCE'], data['WAGE'], '+', label='Observed\n data')
     plt.ylabel('Wage')
     plt.yticks((10, 20, 30, 40), size=10)
     plt.xticks(size=10)
@@ -85,8 +89,8 @@ for name, learner in [('', None),
     plt.xlim(ex_min, ex_max)
     plt.ylim(ymax=48)
 
-    plt.suptitle('Wage as a function of years of education and experience',
-                 size=13)
+    plt.suptitle('Wage in function of years of education and experience     .',
+                 size=14)
 
     #-----------------------------------------------------------------------
     # Some learning
@@ -124,8 +128,8 @@ for name, learner in [('', None),
[…]

### 2014-05-18 wage_data.py
Cleaner figures

@@ -42,9 +42,9 @@ for name, learner in [('', None),
                       (' random forest', ensemble.RandomForestRegressor())]:
     #-----------------------------------------------------------------------
     # Plot a simple pair-wise plot
-    plt.figure(figsize=(6, 5))
+    plt.figure(figsize=(6, 4.5))
 
-    ax_2d = plt.axes([.35, .33, .6, .6])
+    ax_2d = plt.axes([.35, .352, .6, .63])
 
     # Add random offset to limit overlap in points
     n_sample = len(data)
@@ -63,14 +63,15 @@ for name, learner in [('', None),
     cb = plt.colorbar()
     # Squeeze the colorbar a bit left
     bb = cb.ax.get_position()
-    bb.x0 -= .03
+    bb.x0 -= .025
+    bb.x1 -= .01
     cb.ax.set_position(bb)
     cb.ax.set_ylabel('Wage')
     for l in cb.ax.get_yticklabels():
         l.update(dict(size=10))
 
     # Side plot as a function of education
-    ax_ed = plt.axes([.11, .33, .2, .6])
+    ax_ed = plt.axes([.11, .352, .2, .63])
     plt.plot(data['WAGE'], data['EDUCATION'], '+', label='Observed\n data')
     plt.ylim(ed_min, ed_max)
     plt.xlim(xmax=48)
@@ -80,7 +81,7 @@ for name, learner in [('', None),
     plt.yticks(size=10)
 
     # Side plot as a function of experience
-    ax_ex = plt.axes([.35, .09, .48, .2])
+    ax_ex = plt.axes([.35, .09, .48, .22])
     plt.plot(data['EXPERIENCE'], data['WAGE'], '+', label='Observed\n data')
     plt.ylabel('Wage')
     plt.yticks((10, 20, 30, 40), size=10)
@@ -88,9 +89,6 @@ for name, learner in [('', None),
     plt.xlabel('Years of experience')
     plt.xlim(ex_min, ex_max)
     plt.ylim(ymax=48)
-
-    plt.suptitle('Wage in function of years of education and experience     .',
-                 size=14)
 
     #-----------------------------------------------------------------------
     # Some learning
@@ -130,6 +128,5 @@ for name, learner in [('', None),
 
         ax_ex.legend(loc=(-.74, -.22), prop=dict(size=12), frameon=False,
                      handletextpad=.3, labelspacing=1)
[…]

### 2014-05-18 article.tex
Cleaner figures
Iter on figure
Iter on paper, and shorten biblio

--- 2014-05-18 article.tex
@@
 %
-Bridging the gap between machine learning research and applications
+Scikit-learn was born from the observation that most standard
+machine-learning algorithms were out of reach of the users that could
+most benefit from them: researchers --biologists, climate
+scientists, experimental physicists-- or developers --for web
+services or domain-specific applications.
+%
+Implementations mostly consisted of scattered piece of code to download
+on researchers' web pages, with the notable exception
+statistic-specific environments such as the R language \cite{Rmanual}, or
+the Weka java library \cite{hall2009weka}.
 
-\cite{pedregosa2011}
+Scikit-learn aims to bridge the gap between machine-learning research and
+applications by providing a library, and not an environment, in a
+general-purpose language, relying on domain-agnostic data structures
+\cite{pedregosa2011}. Emphasis is put on quality and ease of use, which
+implies focus on installation issues, documentation, and API design
+\cite{buitinck2013ecml}. For the project to be usable in real-world
+settings, computational performance is also a priority.
 
-Impact on education (documentation)
-
-Striving for ease-of-use and quality
-
-Importance of API design \cite{buitinck2013ecml}
-
-\paragraph{The python data ecosystem}
+\paragraph{The Python data ecosystem}
 %
-In a data analysis pipeline, machine learning is only a tiny bit of the
-pipeline, although it is critical.
-
-Great scientific and numeric ecosystem
-\cite{oliphant2007python,varoquaux2013scipy}, but also text processing,
-webservers
-
-NumPy arrays \cite{vanderwalt2011}, SciPy, matplotlib. Cython
-\cite{behnel2011cython}
+Machine learning is only a small part of a data-analysis pipeline, and
+scikit-learn nicely dovetails into the rich Python ecosystem: scientific
+and numeric tools \cite{oliphant2007python,varoquaux2013scipy}, but also
+text processing tools, web servers...
+%
+For its numerical needs, scikit-learn leverages NumPy arrays
+\cite{vanderwalt2011} --data structures for efficient numerical
+computation--, SciPy --a collection of classic numerical algorithms--,
+matplotlib \cite{hunter2007matplotlib} --for scientific plotting-- and
+Cython, to generate and call compiled code in a Python-like syntax
+\cite{behnel2011cython}.
 
@@
 overfitting, model complexity
+
+\begin{figure}
+    %\hspace*{-.015\linewidth}%
+    \includegraphics[width=1.05\linewidth]{wage_data}%
+
+    \caption{Wage as a function of years of experience and education,
+    data from \cite{berndt1991}}
+\end{figure}
+
+\begin{figure}
+    %\hspace*{-.015\linewidth}%
+    \includegraphics[width=1.05\linewidth]{wage_data_random_forest}%
+
+    \caption{Wage prediction from years of experience and education,
+    using random forests}
+\end{figure}
+
+\begin{figure}
+    \hspace*{-.015\linewidth}%
+    \includegraphics[width=1.05\linewidth]{wage_data_linear_svm}%
+
+    \caption{Wage prediction from years of experience and education,
+    using a linear SVM}
+\end{figure}
 
@@
 
+Impact on education (documentation)
+
 \small

### 2014-05-18 article.tex
typo

@@
 Implementations mostly consisted of scattered piece of code to download
-on researchers' web pages, with the notable exception
+on researchers' web pages, with the notable exception of
 statistic-specific environments such as the R language \cite{Rmanual}, or

### 2014-05-18 article.tex
Finish introduction

@@
 \cite{behnel2011cython}.
+%
+A plethora of Python packages can help users with input or preprocessing
+of data, notably Pandas for columnar data \cite{mckinney2012}, scikit-image
+for images, and NLTK for text. Finally, the IPython environment
+\cite{perez2007ipython} is priceless for interactive work.
 
-Pandas for columnar data, scikit-image, NLTK.
-
-IPython for interactive work \cite{perez2007ipython}
-
-\paragraph{Some history of the project}
+\paragraph{Some history}
 %
-PhD project of David Cournapeau and Matthieu Brucher, Parietal-INRIA
-comes along,... 2010 sprint,
-
-LibSVM a good start
-\cite{chang2011libsvm}
+Scikit-learn started circa 2007 as code from David Cournapeau and
+Matthieu Brucher's PhD work, but dwindled down till 2010, year at which the
+Parietal team from INRIA adopted the project, hiring a full time
+engineer. Basics API were defined and an efficient binding of LibSVM
+\cite{chang2011libsvm} gave the project a compelling advantage. In
+December 2011, the first international sprint was organized with generous
+funding from Google. At this point, the project has grown vastly beyond
+INRIA into a worldwide open source effort.
 

### 2014-05-18 article.tex
Formulation

@@
 accessible to all, whether it be in academia or in the industry. It
-fosters the general-purpose Python language, that enjoys a huge growth in
-the scientific world, and well as a striving ecosystem of contributors.
-Here we give a quick introduction to the scikit-learn as well as to
+benefits from 
+the general-purpose Python language, that enjoys a huge growth in
+the scientific world, as well as a striving ecosystem of contributors.
+Here we give a quick introduction to scikit-learn as well as to
 machine-learning basics.

### 2014-05-18 article.tex
Better data description

@@
 \usepackage{authblk}
-\usepackage{dblfloatfix}
-%\usepackage{hyperref}
+%\usepackage{dblfloatfix}
+\usepackage{hyperref}
 \usepackage{graphicx}
@@
 December 2011, the first international sprint was organized with generous
-funding from Google. At this point, the project has grown vastly beyond
+funding from Google. Today, the project has grown vastly beyond
 INRIA into a worldwide open source effort.
@@
 
+\begin{figure}[b]
+    %\hspace*{-.015\linewidth}%
+    \includegraphics[width=1.05\linewidth]{wage_data}%
+
+    \caption{Wage as a function of years of work experience and
+    education; data from \cite{berndt1991}\label{fig:data}.}
+\end{figure}
+
+% Introduce and describe the data
+Machine learning is about extracting rules from data, most often with the
+goal of making decision on new data. As a simple example, we show on
+Fig.\,\ref{fig:data} data about wages in the US, from \cite{berndt1991}:
+11 features, including wage, number of years of education and of work
+experience, on 534 individuals --samples in machine-learning
+jargon. This data
+Code downloading the data to reproduce the figures can be found on
+\url{https://github.com/scikit-learn/SigMobile-paper}.
+%
+This data showcases many of the typical difficulties machine learning.
+The samples are very irregularly distributed, as most individuals
+finished their education after high-school. As a result, they are gapping
+holes in the 2D plot education versus work experience, in which any
+statistical analysis is extrapolation. The data is very noisy: for a
+given education and work experience wages vary widely. This variability
+can probably be explained with missing factors, such as sector of
+activity, but adding them to the analysis creates a more complex picture,
+in 3D or more, rather than 2D, with even more gaps.
+
+% Show an example of learning, first with random forest, and introduce
+% overfit, then with svms, and discuss model complexity and use linear
+% models to link to convention statistics. Conclude with the difference
+% with 1D and 2D, and introduce the curse of dimensionality.
 Discuss what we can learn from the wages examples. Introduce concepts of
@@
 
-\begin{figure}
-    %\hspace*{-.015\linewidth}%
-    \includegraphics[width=1.05\linewidth]{wage_data}%
-
-    \caption{Wage as a function of years of experience and education,
-    data from \cite{berndt1991}}
-\end{figure}
-
-\begin{figure}
+\begin{figure}[b]
     %\hspace*{-.015\linewidth}%
@@
 
-    \caption{Wage prediction from years of experience and education,
-    using random forests}
+    \caption{Wage prediction from years of work experience and education,
+    using random forests.}
 \end{figure}
 
-\begin{figure}
+\begin{figure}[b]
     \hspace*{-.015\linewidth}%
@@
 
-    \caption{Wage prediction from years of experience and education,
-    using a linear SVM}
+    \caption{Wage prediction from years of work experience and education,
+    using a linear SVM.}
 \end{figure}

### 2014-05-18 wage_data.py

@@ -52,8 +52,6 @@ for name, learner in [('', None),
                 data['EDUCATION'] -.25 + .5 * np.random.random(n_sample),
                 c=data['WAGE'], cmap=plt.cm.Blues, vmax=wage_max)
 
-    plt.xlabel('Years of work experience')
-    plt.ylabel('Years of education')
     plt.yticks(size=10)
     plt.xticks(size=10)
 
@@ -86,7 +84,7 @@ for name, learner in [('', None),
     plt.ylabel('Wage')
     plt.yticks((10, 20, 30, 40), size=10)
     plt.xticks(size=10)
-    plt.xlabel('Years of experience')
+    plt.xlabel('Years of work experience')
     plt.xlim(ex_min, ex_max)
     plt.ylim(ymax=48)
 

### 2014-05-18 article.tex
Iter

@@
 
+On Fig.\,\ref{fig:random_forest}, we use a popular machine-learning
+algorithm, Random Forests, to try to predict the wage from the years of
+education, or the years of work experience, either as separate features,
+or combined.
+
 % Show an example of learning, first with random forest, and introduce
@@
     \caption{Wage prediction from years of work experience and education,
-    using random forests.}
+    using random forests.\label{fig:random_forest}}
 \end{figure}

### 2014-05-18 article.tex
Finish intro to ML

@@
 Machine learning is about extracting rules from data, most often with the
-goal of making decision on new data. As a simple example, we show on
+goal of making decision on new data \cite{elemstatlearn}. As a simple example, we show on
 Fig.\,\ref{fig:data} data about wages in the US, from \cite{berndt1991}:
@@
 education, or the years of work experience, either as separate features,
-or combined.
-
-% Show an example of learning, first with random forest, and introduce
-% overfit, then with svms, and discuss model complexity and use linear
-% models to link to convention statistics. Conclude with the difference
-% with 1D and 2D, and introduce the curse of dimensionality.
-Discuss what we can learn from the wages examples. Introduce concepts of
-overfitting, model complexity
+or combined. The patchy square appearance of the prediction is due to the
+inners of the algorithm, and it illustrates the corresponding
+extrapolation mechanism. The predictions fit the data very; maybe too
+well, as, on the prediction solely from the years of education (left of
+the figure), it is hard to believe in the bump at 5 years. This bump is
+probably a classic case of \emph{overfit}: the algorithm is learning its
+prediction from noise in the data. As a result, the prediction error on
+new data will be significantly different than that measured on the
+training data.
 
@@
 
+% then with svms, and discuss model complexity and use linear
+% models to link to convention statistics. Conclude with the difference
+% with 1D and 2D, and introduce the curse of dimensionality.
+%overfitting, model complexity
+
+On Fig.\,\ref{fig:linear_svm}, we use another popular machine-learning
+algorithm, linear Support Vector Machines (SVMs). Unlike random forests,
+it can only learn decisions made with linear functions. It is said that
+it has a lower \emph{model complexity}, because the number of parameters
+to learn from the data is much smaller. The risk here is to
+\emph{underfit}: not fully use the richness of the data, that may not be
+following a linear law. The art of machine learning consists in choosing
+the right family of algorithms (or models) to describe the data well and
+find the sweet spot between under and over fit. As the number of features
+describing the data grows, the number of parameters to learn grows, and
+thus the risk of overfit increases. This core difficulty is known as the
+\emph{curse of dimensionality}.
+
 \begin{figure}[b]
@@
     \caption{Wage prediction from years of work experience and education,
-    using a linear SVM.}
+    using a linear SVM.\label{fig:linear_svm}}
 \end{figure}

### 2014-05-18 article.tex
Iter on the paper

@@
 
-\paragraph{The data matrix}
+\paragraph{Setting up models and feeding them data}
 %
-Blahblah
+Models in scikit-learn, called \emph{estimators}, are objects
+instantiated with all parameters that describe them, but with data. They
+have a {\tt fit} method, that accepts an $(n \times p)$ data
+matrix--numpy array or scipy sparse matrix--, where $n$ is the number of
+samples, and $p$ the number of features--, and possibly a quantity to
+predict, also an array, 1D unless the prediction is \emph{multi-output},
+with floats for a regression task, as in the wages prediction, or
+integers for classification, that is predicting for each sample to which
+class it belongs. The data seldom come in this form, and they must be
+adapted to be suitable for consumption by estimators.
 
@@
 %
-Blahblah
+The goal of supervised models is prediction; they
 

### 2014-05-19 article.tex
iter

--- 2014-05-18 article.tex
@@
 
-\paragraph{The data matrix}
+\paragraph{Setting up models and feeding them data}
 %
-Blahblah
+Models in scikit-learn, called \emph{estimators}, are objects
+instantiated with all parameters that describe them, but with data. They
+have a {\tt fit} method, that accepts an $(n \times p)$ data matrix,
+$\mathbf{X}$ --numpy array or scipy sparse matrix--, where $n$ is the
+number of samples, and $p$ the number of features--, and possibly a
+quantity to predict, $\mathbf{y}$, also an array, 1D unless the
+prediction is \emph{multi-output}, with floats for a regression task, as
+in the wages prediction, or integers for classification, that is
+predicting for each sample to which class it belongs. The data seldom
+come in this form, and they must be adapted to be suitable for
+consumption by estimators.
 
@@
 %
-Blahblah
+The goal of supervised models is prediction. The corresponding estimators
+have a
 

### 2014-05-19 article.tex
Iter on paper

@@
 The goal of supervised models is prediction. The corresponding estimators
-have a
+have a {\tt predict} method, that takes a data matrix $\mathbf{X}$ and
+returns a predicted $\mathbf{y}$.
 
@@
 %
-Model evaluation:
-cross-validation: the concept, cross-val iterators, and cross\_val\_score
-
-Problem of multiple metrics.
+The data used to train a model can be used to measure the prediction
+error, as it would be enable to distinguish noise from signal, and thus
+overly optimistic. The correct way is to leave out \emph{test} data,
+untouched during training. Often a \emph{cross-validation} scheme is
+used, where the data is repeatedly split into \emph{train} and
+\emph{test} subsets. Scikit-learn provides a complete framework for
+cross-validation. Specific iterators are used to describe the subsets and
+several functions and objects accept as arguments these iterators to
+perform cross-validation internally. For instance the {\tt
+cross\_val\_score} function measures the prediction score of an
+estimator.
 

### 2014-05-19 article.tex
fixes for NLP section

@@
 In the NLP application domain,
-features are typically 
-where features are typically frequencies or boolean indicators
-of very large numbers of rare events, such as
-``document contains the term \textsf{v1agr@}'' (spam filtering)
-or ``next token matches the Roman numeral regex'' (named-entity recognition).
-To ensure scalability in the face of such data,
+features are typically frequencies or boolean indicators
+of very large numbers of rare events.
+In document classification tasks such spam filtering,
+these events may be booleans indicating
+``document contains the term \textsf{v1agr@}''
+(and so on for all words encountered in the training set).
+In other NLP problems such as finding proper names and their classes (NER),
+events such as ``next token matches the Roman numeral regex'' may be used.
+To ensure scalability in the face of such extremely high-dimensional data,
 SciPy's sparse matrix data structures are supported by many estimators.\footnote{
@@
 and finally trains the model.
+The feature extractor (vectorizer) takes text as strings
+and produces a sparse matrix of weights based on frequencies.
 The parameters for the feature extractor and the LogReg classifier
@@
 Does this classifier work?
-We can find out by classifying some of our own movie reviews.\footnote{
+We can find out by classifying two of our own movie reviews.\footnote{
   The code can be tried interactively by simply copy-pasting it

### 2014-05-19 article.tex
tinyurl for movie review dataset

@@
 
-temp, \_ = urlretrieve('http://www.cs.cornell.edu/people/pabo/movie-review-data/review\_polarity.tar.gz')
+temp, \_ = urlretrieve('http://preview.tinyurl.com/review-tgz')
 with tarfile.open(temp) as tar:

### 2014-05-19 article.tex
minted -> listings
Needs tweaking (font too big)

@@
 \usepackage{graphicx}
-\usepackage{minted}
+\usepackage{listings}
 \usepackage[numbers,square]{natbib}
@@
 
-\definecolor{rulecolor}{rgb}{0.80,0.80,0.80}
-\newminted{python}{frame=single,rulecolor=\color{rulecolor}}
-
 \title{Scikit-learn: machine learning without learning the machinery}
@@
 \begin{document}
+\lstset{language=Python}
 
@@
 % the tarball from Pang et al. contain top-level cruft.
-\begin{pythoncode}
+\begin{lstlisting}
 import os.path
@@
 from urllib import urlretrieve
-from sklearn.datasets import load\_files
-from sklearn.feature\_extraction.text import TfidfVectorizer
-from sklearn.linear\_models import LogisticRegression
-from sklearn.pipeline import make\_pipeline
+from sklearn.datasets import load_files
+from sklearn.feature_extraction.text import TfidfVectorizer
+from sklearn.linear_models import LogisticRegression
+from sklearn.pipeline import make_pipeline
 
-temp, \_ = urlretrieve('http://preview.tinyurl.com/review-tgz')
+temp, _ = urlretrieve('http://preview.tinyurl.com/review-tgz')
 with tarfile.open(temp) as tar:
-    tar.extractall(path='movie\_reviews')
-data = load\_files(os.path.join('movie\_reviews', 'txt\_sentoken'))
+    tar.extractall(path='movie_reviews')
+data = load_files(os.path.join('movie_reviews', 'txt_sentoken'))
 
-clf = make\_pipeline(TfidfVectorizer(min\_df=2, dtype=float,
-                                    sublinear\_tf=True, ngram\_range=(1, 2),
-                                    strip\_accents='unicode'),
-                     LogisticRegression(random\_state=623, C=5000))
+clf = make_pipeline(TfidfVectorizer(min_df=2, dtype=float,
+                                    sublinear_tf=True, ngram_range=(1, 2),
+                                    strip_accents='unicode'),
+                     LogisticRegression(random_state=623, C=5000))
 clf.fit(data.data, data.target)
-\end{pythoncode}
+\end{lstlisting}
 
@@
 
-\begin{pythoncode}
->>> clf.predict\_proba(["This movie is the worst I ever saw.",
+\begin{lstlisting}
+>>> clf.predict_proba(["This movie is the worst I ever saw.",
 ...                    "Shawshank Redemption, eat your heart out!"])
@@
        [  1.68757754e-01,   8.31242246e-01]])
-\end{pythoncode}
+\end{lstlisting}
 
@@
 
-\begin{pythoncode}
->>> from sklearn.cross\_validation import cross\_val\_score
->>> score = cross\_val\_score(clf, data.data, data.target, cv=5)
+\begin{lstlisting}
+>>> from sklearn.cross_validation import cross_val_score
+>>> score = cross_val_score(clf, data.data, data.target, cv=5)
 >>> score
@@
 0.88200000000000001
-\end{pythoncode}
+\end{lstlisting}
 

### 2014-05-21 article.tex
Finish the unsupervised learning part

@@
 cross\_val\_score} function measures the prediction score of an
-estimator.
-
-CV objects: GridSearchCV and RandomSearchCV, as well as FooBarCV objects.
+estimator. Cross-validation can also be used to tuned the meta-parameters
+of an estimator, such as the sparsity of a sparse model. For this,
+specific estimators, \emph{meta-estimators}, such as the {\tt
+GridSearchCV}, take another estimator at construction time, and use
+cross-validation internally to set its parameter. When used together with
+{\tt cross\_val\_score}, they perform \emph{nested} cross-validation,
+\emph{i.e.}~the meta-parameters are set independently of the test data?
 
@@
 %
-Unsupervised learning: a variety of usage pattern.
-
-Many things can be seen as Transformers
-
-Introduce the pipeline
+Unsupervised learning covers all learning applications in which there is
+no clearly identified variable to predict. For instance, in a clustering
+application, the task is to group together observations that are similar.
+Dimension reduction tries to find simplified representations that capture
+well the properties of the data. Novelty detection finds in a new dataset
+observations that differ from the data in the train set. As the uses of
+unsupervised learning are very diverse, it is not possible to have an API
+as uniform as for supervised learning. While some unsupervised
+estimators can be used to predict a characteristic of new data, such as
+in novelty detection, many are useful to transform data, as in dimension
+random. Scikit-learn estimators can have a {\tt transform} method, used
+for this purpose. A {\tt Pipeline} estimator can then chain estimators to
+form a new estimator that applies transformations before calling the {\tt
+fit} or {\tt predict} method of the last object.
 

### 2014-05-21 article.tex
Compress code example

@@
 \usepackage{magazine-style}
+
+% Listing settings for Python
+\lstset{language=python,
+	extendedchars=true,
+	xleftmargin = 0pt,
+	%frame=single,
+	%rulecolor=\color{lightgrey},
+        aboveskip = 0.5ex,
+        belowskip = 0.6ex,
+	basicstyle=\small\sffamily,
+	escapebegin={\color{green}},
+        keywordstyle=\bfseries\color{blue!50!black},
+	%identifierstyle=\sffamily\small,
+	commentstyle=\color{green!40!black},
+	stringstyle=\rmfamily\color{red!60!black},
+	showstringspaces=false,
+	tabsize=2,
+	breaklines=true,
+	%classoffset=1,
+        morekeywords={{,},=,:}, %keywordstyle=\color{SkoleYellow},
+	%classoffset=0,
+	backgroundcolor=\color{green!10},
+	fillcolor=\color{green!10},
+}
 
@@
 
+\begin{figure*}
+\lstinputlisting[lastline=12]{nlp_example.py}
+\end{figure*}
+
 The following example script is a fully functioning
@@
 % the tarball from Pang et al. contain top-level cruft.
-\begin{lstlisting}
-import os.path
-import tarfile
-from urllib import urlretrieve
-from sklearn.datasets import load_files
-from sklearn.feature_extraction.text import TfidfVectorizer
-from sklearn.linear_models import LogisticRegression
-from sklearn.pipeline import make_pipeline
-
-temp, _ = urlretrieve('http://preview.tinyurl.com/review-tgz')
-with tarfile.open(temp) as tar:
-    tar.extractall(path='movie_reviews')
-data = load_files(os.path.join('movie_reviews', 'txt_sentoken'))
-
-clf = make_pipeline(TfidfVectorizer(min_df=2, dtype=float,
-                                    sublinear_tf=True, ngram_range=(1, 2),
-                                    strip_accents='unicode'),
-                     LogisticRegression(random_state=623, C=5000))
-clf.fit(data.data, data.target)
-\end{lstlisting}
+%\begin{lstlisting}
+%import os.path
+%import tarfile
+%from urllib import urlretrieve
+%from sklearn.datasets import load_files
+%from sklearn.feature_extraction.text import TfidfVectorizer
+%from sklearn.linear_models import LogisticRegression
+%from sklearn.pipeline import make_pipeline
+%
+%temp, _ = urlretrieve('http://preview.tinyurl.com/review-tgz')
+%with tarfile.open(temp) as tar:
+%    tar.extractall(path='movie_reviews')
+%data = load_files(os.path.join('movie_reviews', 'txt_sentoken'))
+%
+%clf = make_pipeline(TfidfVectorizer(min_df=2, dtype=float,
+%                                    sublinear_tf=True, ngram_range=(1, 2),
+%                                    strip_accents='unicode'),
+%                     LogisticRegression(random_state=623, C=5000))
+%clf.fit(data.data, data.target)
+%\end{lstlisting}
 

### 2014-05-21 nlp_example.py

@@ -1,16 +1,13 @@
 import urllib, tarfile  # Standard library imports
 from sklearn import linear_model, pipeline, datasets, feature_extraction
 
-urllib.urlretrieve('http://www.cs.cornell.edu/people/pabo/movie-review-data/review_polarity.tar.gz', 'tmp.tar.gz')
-with tarfile.open('tmp.tar.gz') as tar:
-    tar.extractall(path='.')
+urllib.urlretrieve('http://bit.ly/1juuXIr', 'tmp.tgz')
+tarfile.open('tmp.tgz').extractall(path='.')
 data = datasets.load_files('txt_sentoken')
 
-clf = pipeline.make_pipeline(feature_extraction.text.TfidfVectorizer(min_df=2,
-                                    dtype=float,
-                                    sublinear_tf=True, ngram_range=(1, 2),
-                                    strip_accents='unicode'),
-                     linear_model.LogisticRegression(random_state=623, C=5000))
+clf = pipeline.Pipeline([('step1', feature_extraction.text.TfidfVectorizer(min_df=2,
+                                                 sublinear_tf=True, ngram_range=(1, 2))),
+                         ('step2', linear_model.LogisticRegression(C=5000))])
 clf.fit(data.data, data.target)
 
 # Demo
@@ -21,4 +18,5 @@ clf.predict_proba(["Shawshank Redemption, eat your heart out!"])[0, 1]  # positi
 
 # Measure the prediction score by cross validation
 from sklearn import cross_validation
-cross_validation.cross_val_score(clf, data.data, data.target, cv=5)
+print cross_validation.cross_val_score(clf, data.data, data.target, cv=5,
+verbose=10)

### 2014-05-21 article.tex
Compress a but more the code

@@
 \begin{lstlisting}
->>> clf.predict_proba(["This movie is the worst I ever saw.",
-...                    "Shawshank Redemption, eat your heart out!"])
-array([[  9.99293000e-01,   7.07000280e-04],
-       [  1.68757754e-01,   8.31242246e-01]])
+>>> clf.predict(["Worst movie I ever saw",
+...      "Godzilla, eat your heart out!"])
+array([0, 1])
 \end{lstlisting}
@@
 \begin{lstlisting}
->>> from sklearn.cross_validation import cross_val_score
->>> score = cross_val_score(clf, data.data, data.target, cv=5)
->>> score
-array([ 0.91  ,  0.8825,  0.88  ,  0.8775,  0.86  ])
->>> score.mean()
-0.88200000000000001
+>>> from sklearn import cross_validation
+>>> cross_validation.cross_val_score(clf, data.data, data.target)
+array([ 0.91, 0.8825, 0.88, 0.8775, 0.86])
 \end{lstlisting}

### 2014-05-21 nlp_example.py

@@ -11,12 +11,13 @@ clf = pipeline.Pipeline([('step1', feature_extraction.text.TfidfVectorizer(min_d
 clf.fit(data.data, data.target)
 
 # Demo
-clf.predict_proba(["This movie is the worst I ever saw."])[0, 0]  # negative probability
+clf.predict_proba(["Worst movie I ever saw"])[0, 0]  # negative probability
 # 0.99929299972040175
-clf.predict_proba(["Shawshank Redemption, eat your heart out!"])[0, 1]  # positive probability
-# 0.83124224591605844
+clf.predict_proba(["Godzilla, eat your heart out!"])[0, 1]  # positive probability
+# 0.64533559029803611
+print clf.predict(["This movie is the worst I ever saw.",
+"Shawshank Redemption, eat your heart out!"])
 
 # Measure the prediction score by cross validation
 from sklearn import cross_validation
-print cross_validation.cross_val_score(clf, data.data, data.target, cv=5,
-verbose=10)
+print cross_validation.cross_val_score(clf, data.data, data.target)

### 2014-05-21 nlp_example.py
Annotate the code

@@ -1,9 +1,9 @@
 import urllib, tarfile  # Standard library imports
 from sklearn import linear_model, pipeline, datasets, feature_extraction
 
-urllib.urlretrieve('http://bit.ly/1juuXIr', 'tmp.tgz')
-tarfile.open('tmp.tgz').extractall(path='.')
-data = datasets.load_files('txt_sentoken')
+urllib.urlretrieve('http://bit.ly/1juuXIr', 'tmp.tgz') # Download the data
+tarfile.open('tmp.tgz').extractall(path='.') # Extract it
+data = datasets.load_files('txt_sentoken') # And load it
 
 clf = pipeline.Pipeline([('step1', feature_extraction.text.TfidfVectorizer(min_df=2,
                                                  sublinear_tf=True, ngram_range=(1, 2))),

### 2014-05-21 article.tex
Annotate the code
Add forgotten file
Place the code float at the bottom, close to where it is necessary
Compress

@@
 \usepackage{authblk}
-%\usepackage{dblfloatfix}
+\usepackage{dblfloatfix}
 \usepackage{hyperref}
@@
 
-\paragraph{Project vision}
+\paragraph{Project vision.}
 %
@@
 
-\paragraph{The Python data ecosystem}
+\paragraph{The Python data ecosystem.}
 %
@@
 
-\paragraph{Some history}
+\paragraph{Some history.}
 %
@@
 
-\paragraph{Setting up models and feeding them data}
+\paragraph{Setting up models and feeding them data.}
 %
@@
 
-\paragraph{Supervised models: learning to predict}
+\paragraph{Supervised models: learning to predict.}
 %
@@
 
-\paragraph{Model evaluation and parameter selection}
+\begin{figure*}[b]
+% To have the float on the bottom, we need this declared on the page
+% before where it is going to be printed
+\lstinputlisting[lastline=12]{nlp_example.py}
+\caption{Simple text-processing code to download movie reviews from
+Internet and learn positive versus negative ratings.\label{fig:code}}
+\end{figure*}
+
+\paragraph{Model evaluation and parameter selection.}
 %
@@
 
-\paragraph{Unsupervised models: learning to transform}
+\paragraph{Unsupervised models: learning to transform.}
 %
@@
 
-\section{In practice}
+\section{In practice: putting scikit-learn to good use}
 
-\paragraph{A simple text-mining example}
+\paragraph{A simple text-mining example.}
 %
-% Lars: the following text presumes that NumPy, estimators, and the
-% X/y convention for supervised learning have already been introduced.
+Not every dataset comes as a set of numbers. For instance in natural
+language processing (NLP), each observation is a text, \emph{i.e.} a
+string, that needs to be transformed into the data matrix expected by
+estimators. Typical, features used to build decisions indicate the
+presence of certain words, and often their frequencies.
+In spam filtering, it can very relevant to extract from a document that
+it contains the term \textsf{viagra}, and possibly how many times.
+In other NLP problems such as finding proper names,
+events such as ``next token is the Roman numeral'' may be relevant, as in
+``J.P. Kennedy III''. Counting bigrams, or $n$-grams, can be a way of
+capturing this information. Reusing the {\tt transform} API, scikit-learn
+provide a few \emph{feature extraction} objects, that can turn such
+unstructured data into a data matrix.
+The $\textbf{X}$ matrix is in these situations a
+matrix of counting statistics. It is full of zeros, in the sens that very
+often a word is present only in a few documents. For better scalability,
+it can be encoded using SciPy's sparse matrices, that many scikit-learn
+estimators process efficiently.
 
-Scikit-learn was designed around NumPy principles and data structures,
-so virtually all estimators take arrays of numbers as input
-and produce arrays as output.
-That doesn't mean that it is limited to traditional number crunching domains,
-though.
-The library has been successfully employed in text classification
-and natural language processing problems,
-and utility code is provided to translate the unstructured, symbolic data
-encountered in these domains
-to the rigid matrix/vector structure that estimators expect.
-
-In the NLP application domain,
-features are typically frequencies or boolean indicators
-of very large numbers of rare events.
-In document classification tasks such spam filtering,
-these events may be booleans indicating
-``document contains the term \textsf{v1agr@}''
-(and so on for all words encountered in the training set).
-In other NLP problems such as finding proper names and their classes (NER),
-events such as ``next token matches the Roman numeral regex'' may be used.
-To ensure scalability in the face of such extremely high-dimensional data,
-SciPy's sparse matrix data structures are supported by many estimators.\footnote{
-  Support for these is added on a by-need basis,
-  because using sparse data structures puts constraints on algorithm design
-  and implementation.
-}
-
-\begin{figure*}
-\lstinputlisting[lastline=12]{nlp_example.py}
-\end{figure*}
-
-The following example script is a fully functioning
-sentiment polarity classifier for movie reviews:
-given a review, it outputs the probability that the review is positive
-about the movie it concerns.
-The script demonstrates the use of pipelines,
-and also shows how machine learning ties into Python's
-``batteries included'' philosophy:
-downloading, unpacking and learning from a dataset
-are all done in one programming language.
-
-% XXX make_pipeline is not in 0.14.1, we should release 0.15 sometime
-% The example can be made a bit shorter by skipping the movie_reviews dir;
-% the tarball from Pang et al. contain top-level cruft.
-%\begin{lstlisting}
-%import os.path
-%import tarfile
-%from urllib import urlretrieve
-%from sklearn.datasets import load_files
-%from sklearn.feature_extraction.text import TfidfVectorizer
-%from sklearn.linear_models import LogisticRegression
-%from sklearn.pipeline import make_pipeline
-%
-%temp, _ = urlretrieve('http://preview.tinyurl.com/review-tgz')
-%with tarfile.open(temp) as tar:
-%    tar.extractall(path='movie_reviews')
-%data = load_files(os.path.join('movie_reviews', 'txt_sentoken'))
-%
-%clf = make_pipeline(TfidfVectorizer(min_df=2, dtype=float,
-%                                    sublinear_tf=True, ngram_range=(1, 2),
-%                                    strip_accents='unicode'),
-%                     LogisticRegression(random_state=623, C=5000))
-%clf.fit(data.data, data.target)
-%\end{lstlisting}
+On Fig.\,\ref{fig:code} we give an example of fully functioning code to
+do sentiment polarity classification for movie reviews: given a review,
+it outputs whether the review is positive about the movie it concerns.
+The code demonstrates the use of pipelines, and also shows how machine
+learning ties into unrelated Python's modules: downloading, unpacking and
+learning from a dataset are all done in one programming language.
 
@@
 \cite{pang2004} and unpacks it to disk using standard Python libraries.
-It then loads these into memory using the scikit-learn function
-\texttt{load\_files}, makes a pipeline of a \textsf{tf--idf} feature extractor
-\cite{rennie2003tackling} and logistic regression,
-and finally trains the model.
-The feature extractor (vectorizer) takes text as strings
-and produces a sparse matrix of weights based on frequencies.
-The parameters for the feature extractor and the LogReg classifier
-were tuned offline; the random-number generator seed is set by hand
-to get reproducible results.
+It then loads these using scikit-learn data loading functions, makes a
+pipeline of a \textsf{tf--idf} feature extractor
+\cite{rennie2003tackling} --than turns strings into a sparse matrix based
+on word frequencies-- and logistic regression, and finally trains the
+model. The parameters for the feature extractor and the LogReg classifier
+were set manually.
 
@@
 
-\small
 \bibliography{paper}

### 2014-05-21 article.tex
Finish pass on text example

@@
 provide a few \emph{feature extraction} objects, that can turn such
-unstructured data into a data matrix.
-The $\textbf{X}$ matrix is in these situations a
+unstructured data into a data matrix. In these situations, the
+$\textbf{X}$ matrix is a
 matrix of counting statistics. It is full of zeros, in the sens that very
@@
 
-Does this classifier work?
-We can find out by classifying two of our own movie reviews.\footnote{
+We can test our classifier on two new movie reviews.\footnote{
   The code can be tried interactively by simply copy-pasting it
@@
 }
-
 \begin{lstlisting}
@@
 
-Here, we asked the model to predict the probability of a review being positive
-for a batch of two reviews. It returns the negative probability in the first
-column, the positive probability in the second.
-As we can see, the first review gets a 99.9\% probability for being negative,
-while the second has an 83\% probability of being positive.
-
-Of course, that's just anecdotal evidence. We can get a figure for the
-classifier's accuracy by retraining it in a five-fold cross-validation scheme:
-
+Here, we asked the estimator to predict if the reviews were positive for
+a batch of two reviews. The first review is predicted has being not
+positive, \emph{i.e.} negative, while the second one is positive. We have
+anecdotal evidence that the classifier work. To get a measure for the
+classifier's accuracy we retraining it in a cross-validation scheme, this
+yield the fraction of correct answers on three different folds:
 \begin{lstlisting}
@@
 >>> cross_validation.cross_val_score(clf, data.data, data.target)
-array([ 0.91, 0.8825, 0.88, 0.8775, 0.86])
+array([0.9026946, 0.8723724, 0.8633634])
 \end{lstlisting}

### 2014-05-21 article.tex
Intro to big data

@@
 %
+``Big data'' is the latest trend of data analytics: applying machine
+learning to large datasets can extract information so far unexploited.
+These large datasets present of course new technical challenges, in
+particular to be computationally tractable. It is important to
+distinguish the case of very large number of samples, or very large
+number of features. For a very large number of features, on-the-fly data
+reductions, that exploit latent redundancy in the features, come in
+handy. In scikit-learn, one could for instance resort to
+random projections, that exploit the Johnson-Lindenstrauss lemma to give
+good approximations of Euclidean distance in the datasets. In the
+situation of a very large number of samples, some estimator expose
+\emph{on-line} algorithms: they can be fit incrementally, with successive
+calls to their {\tt partial\_fit} method, and converge in a small number
+of passes on the data. To alleviate one of the main practical challenges of
+big data, data-transfer latency, optimal data-fetching code can be used
+in parallel with the calls to {\tt partial\_fit}.
+
 Partial fit, random projections, HashingVectorizer

### 2014-05-22 article.tex
Introduce the HashingVectorizer

@@
 random projections, that exploit the Johnson-Lindenstrauss lemma to give
-good approximations of Euclidean distance in the datasets. In the
+good approximations of Euclidean distances in the dataset. In the
 situation of a very large number of samples, some estimator expose
@@
 
-Partial fit, random projections, HashingVectorizer
+Dealing with a large corpus of text, faces both challenges: many samples,
+and a very high dimensional space. In addition, the dimensionality
+tends to increase with sample size, as the number of different words used 
+is greater in large corpuses. Beyond the mere size of the data matrix
+created by a standard vectorization approach, as above, another challenge
+is that the vectorization must be done in two passes over the documents,
+as the size of the output vector is not known before hand. Here, hashing,
+a trick similar to random projection, can be used to tackle both of these
+challenge and produce a streaming vectorizer, the {\tt HashingVectorizer}
+in scikit-learn. The idea is to use a hash table, storing hashes of the
+words, rather than the words themselves. The dimensionality of the hash
+table is known from the hash size.
 

### 2014-05-22 article.tex
Iter on paper

@@
 {\tt cross\_val\_score}, they perform \emph{nested} cross-validation,
-\emph{i.e.}~the meta-parameters are set independently of the test data?
+\emph{i.e.}~the meta-parameters are set independently of the test data.
 
@@
 Not every dataset comes as a set of numbers. For instance in natural
-language processing (NLP), each observation is a text, \emph{i.e.} a
+language processing (NLP), each observation is a text, \emph{i.e.}~a
 string, that needs to be transformed into the data matrix expected by
@@
 it contains the term \textsf{viagra}, and possibly how many times.
-In other NLP problems such as finding proper names,
+In other NLP problems, such as finding proper names,
 events such as ``next token is the Roman numeral'' may be relevant, as in
-``J.P. Kennedy III''. Counting bigrams, or $n$-grams, can be a way of
-capturing this information. Reusing the {\tt transform} API, scikit-learn
-provide a few \emph{feature extraction} objects, that can turn such
+``J.P. Kennedy III''. A way to capture this information is to count bigrams,
+or $n$-grams. Reusing the {\tt transform} API, scikit-learn
+provides \emph{feature extraction} objects, that  such
 unstructured data into a data matrix. In these situations, the
 $\textbf{X}$ matrix is a
-matrix of counting statistics. It is full of zeros, in the sens that very
+matrix of counting statistics. It is full of zeros, in the sens that
 often a word is present only in a few documents. For better scalability,
@@
 learning from a dataset are all done in one programming language.
-
 This script first fetches the hand-labeled movie review dataset from
@@
 on word frequencies-- and logistic regression, and finally trains the
-model. The parameters for the feature extractor and the LogReg classifier
+model. Parameters for the feature extractor and the LogReg classifier
 were set manually.
@@
 classifier's accuracy we retraining it in a cross-validation scheme, this
-yield the fraction of correct answers on three different folds:
+yields the fraction of correct answers on three different folds:
 \begin{lstlisting}

### 2014-05-22 article.tex
Iteration/compression

@@
 experience, on 534 individuals --samples in machine-learning
-jargon. This data
+jargon. 
 Code downloading the data to reproduce the figures can be found on
@@
 %
-This data showcases many of the typical difficulties machine learning.
-The samples are very irregularly distributed, as most individuals
-finished their education after high-school. As a result, they are gapping
-holes in the 2D plot education versus work experience, in which any
+This data showcases many of the typical difficulties in machine learning.
+Samples are very irregularly distributed, as most individuals
+finished studies after high-school. As a result, there are gapping
+holes in the 2D plot of education versus work experience, in which any
 statistical analysis is extrapolation. The data is very noisy: for a
 given education and work experience wages vary widely. This variability
-can probably be explained with missing factors, such as sector of
+is probably explained by missing factors, such as sector of
 activity, but adding them to the analysis creates a more complex picture,
@@
 On Fig.\,\ref{fig:random_forest}, we use a popular machine-learning
-algorithm, Random Forests, to try to predict the wage from the years of
+algorithm, Random Forests, to predict the wage from the years of
 education, or the years of work experience, either as separate features,
@@
 algorithm, linear Support Vector Machines (SVMs). Unlike random forests,
-it can only learn decisions made with linear functions. It is said that
+it learns decisions with linear functions. It is said that
 it has a lower \emph{model complexity}, because the number of parameters
@@
 learning to large datasets can extract information so far unexploited.
-These large datasets present of course new technical challenges, in
+These large datasets present new technical challenges, in
 particular to be computationally tractable. It is important to
@@
 reductions, that exploit latent redundancy in the features, come in
-handy. In scikit-learn, one could for instance resort to
+handy. In scikit-learn, one could for instance use
 random projections, that exploit the Johnson-Lindenstrauss lemma to give
-good approximations of Euclidean distances in the dataset. In the
-situation of a very large number of samples, some estimator expose
+good approximations of Euclidean distances in the dataset. For
+a very large number of samples, some estimators expose
 \emph{on-line} algorithms: they can be fit incrementally, with successive
@@
 
-Dealing with a large corpus of text, faces both challenges: many samples,
+Dealing with a large corpus of text faces both challenges: many samples,
 and a very high dimensional space. In addition, the dimensionality

### 2014-06-22 article.tex
iter

@@
 
-Goal: enable anybody to contribute, have a controlled process, grow.
+The greatest asset of the scikit-learn project is its breadth of
+contributors, coming from different backgrounds, working in different
+institutions, with a variety of applications --xxx contributors total,
+xxx active core contributors in the past year--. Indeed, not only is
+there strength in numbers, but also the convergence of points of views
+make better code and better design. From a developer's perspective, the
+project thus is managed with the goals of enabling newcomers to
+contribute with a tightly controlled process.
 

### 2014-06-30 article.tex
Iter paper

--- 2014-05-22 article.tex
@@
 
-Goal: enable anybody to contribute, have a controlled process, grow.
+The greatest asset of the scikit-learn project is its breadth of
+contributors, coming from different backgrounds, working in different
+institutions, with a variety of applications --more than 200 contributors
+total, 15 active core contributors in the past year. Indeed, not only
+is there strength in numbers, but also the convergence of points of views
+make better code and better design. From a developer's perspective, the
+project thus is managed with the goals of enabling newcomers to
+contribute while keeping a high quality assurance. For these purposes, we
+rely heavily on Github --\url{http://github.com}-- with an online
+development workflow relying on distributed version control using git,
+and intensive code review. The code review process is a great asset to
+raising the quality of the code, from all points of view: algorithmic,
+numerical robustness, homogeneity of APIs, ease of use, documentation.
+Another central aspect for quality in the project is the heavy use of
+unit testing (up to 88\% of the lines of code are covered), run
+continuously on all contributions thanks to the travis platform
+(\url{http://travis-ci.org}).
 
-Github, tests (link to travis)
+The open-source, community-driven model of scikit-learn has fantastically
+payed off. However, we find that it does not fit in the standard
+professional incentive structure, whether it be in academia or in the
+industry. Indeed, getting due credit for grass-root, free, software that
+"just works" is difficult. None of the grant proposals submitted to fund
+development by core contributors have so far been accepted, and there is
+no straightforward revenue model. Even more challenging is rewarding
+properly the long tail of small contributors. Finally, active
+contributors tend to be victim of the project's success, and be promoted
+to leading positions which leave them little time.
 
-Difficulty of getting credit, of rewarding properly the long tail of small
-contributors, of finding funding. Problem of brain drain.
-
-Difficulty of project scope (ideally: first cover all of statistical
-learning classics \cite{elemstatlearn}) and default parameters
+Scikit-learn has become a reference machine learning toolkit. As such, it
+should offer a wide variety of learning algorithms. However, each new
+feature, each new line of code comes with a significant maintenance cost.
+Uncontrolled growth would lead the project to pick up weight until it
+comes to a crawl. The difficulty lies in defining the project scope and
+prioritizing methods. We strive to integrate all classical methods that
+are most useful for prediction tasks. We favor time-tested algorithm and
+put great care in the choice of default parameters.
 

### 2014-06-30 article.tex
Finish?

@@
 comes to a crawl. The difficulty lies in defining the project scope and
-prioritizing methods. We strive to integrate all classical methods that
-are most useful for prediction tasks. We favor time-tested algorithm and
-put great care in the choice of default parameters.
+prioritizing methods. We strive to identify and integrate all major
+methods that are most useful for prediction tasks. We favor time-tested
+algorithms and put great care in the choice of default parameters. Our
+secret ambition is that the documentation, that we strive to make
+didactic and pragmatic, can provide an executable alternative to
+textbooks, as all the figures are generated with code examples.
 
-We have found that implementing, even a standard algorithm, really well,
-can require a lot of domain knowledge. Thus it is natural that specific
-libraries span up to solve it. Some adopt scikit-learn API and standard,
-and we hope that scikit-learn has a structural effect on the environment.
-
-Impact on education (documentation)
+%We have found that implementing, even a standard algorithm, really well,
+%can require a lot of domain knowledge. Thus it is natural that specific
+%libraries span up to solve it. Some adopt scikit-learn API and standard,
+%and we hope that scikit-learn has a structural effect on the environment.
 

### 2014-06-30 article.tex
Finish??

@@
 should offer a wide variety of learning algorithms. However, each new
-feature, each new line of code comes with a significant maintenance cost.
+feature, each new line of code, comes with a significant maintenance cost.
 Uncontrolled growth would lead the project to pick up weight until it
-comes to a crawl. The difficulty lies in defining the project scope and
+came to a crawl. The difficulty lies in defining the project scope and
 prioritizing methods. We strive to identify and integrate all major
-methods that are most useful for prediction tasks. We favor time-tested
+methods most useful for prediction tasks. We favor time-tested
 algorithms and put great care in the choice of default parameters. Our
@@
 didactic and pragmatic, can provide an executable alternative to
-textbooks, as all the figures are generated with code examples.
+textbooks, with all the figures generated by code examples.
 

### 2014-07-01 article.tex
typo
copyedit, esp. the NLP example part

@@
 Samples are very irregularly distributed, as most individuals
-finished studies after high-school. As a result, there are gapping
+finished studies after high-school. As a result, there are gaping
 holes in the 2D plot of education versus work experience, in which any
@@
 Not every dataset comes as a set of numbers. For instance in natural
-language processing (NLP), each observation is a text, \emph{i.e.}~a
-string, that needs to be transformed into the data matrix expected by
-estimators. Typical, features used to build decisions indicate the
-presence of certain words, and often their frequencies.
-In spam filtering, it can very relevant to extract from a document that
-it contains the term \textsf{viagra}, and possibly how many times.
+language processing (NLP), observations are strings,
+which needs to be transformed into the data matrix expected by
+estimators. Often, features used to build decisions indicate the
+presence of certain words, or their frequencies.
+In spam filtering, it can be very relevant to extract from a document that
+it contains the term \textsf{viagra}.
 In other NLP problems, such as finding proper names,
-events such as ``next token is the Roman numeral'' may be relevant, as in
-``J.P. Kennedy III''. A way to capture this information is to count bigrams,
-or $n$-grams. Reusing the {\tt transform} API, scikit-learn
-provides \emph{feature extraction} objects, that  such
+events such as ``next token is a Roman numeral'' may be relevant, as in
+``J.P. Kennedy III''.
+Reusing the {\tt transform} API, scikit-learn
+provides \emph{feature extraction} objects, that turn such
 unstructured data into a data matrix. In these situations, the
-$\textbf{X}$ matrix is a
-matrix of counting statistics. It is full of zeros, in the sens that
-often a word is present only in a few documents. For better scalability,
-it can be encoded using SciPy's sparse matrices, that many scikit-learn
-estimators process efficiently.
+$\textbf{X}$ matrix is a matrix of counting statistics.
+Since each word is typically present in only a few documents,
+the data matrix will be mostly zeros.
+SciPy offers a \emph{sparse matrix} package
+that stores the zeros in such matrices implicitly,
+and many scikit-learn estimators use that package to improve scalability.
 
@@
 learning from a dataset are all done in one programming language.
-This script first fetches the hand-labeled movie review dataset from
+This script first fetches a hand-labeled movie review dataset
 \cite{pang2004} and unpacks it to disk using standard Python libraries.
@@
 pipeline of a \textsf{tf--idf} feature extractor
-\cite{rennie2003tackling} --than turns strings into a sparse matrix based
+\cite{rennie2003tackling} --that turns strings into a sparse matrix based
 on word frequencies-- and logistic regression, and finally trains the
-model. Parameters for the feature extractor and the LogReg classifier
+model. Parameters for the feature extractor and the classifier
 were set manually.
@@
 contribute while keeping a high quality assurance. For these purposes, we
-rely heavily on Github --\url{http://github.com}-- with an online
+rely heavily on Github (\url{http://github.com}) with an online
 development workflow relying on distributed version control using git,

### 2014-07-01 article.tex
typo

@@
 anecdotal evidence that the classifier work. To get a measure for the
-classifier's accuracy we retraining it in a cross-validation scheme, this
+classifier's accuracy we retrain it in a cross-validation scheme, this
 yields the fraction of correct answers on three different folds:

### 2014-07-01 article.tex
grammar: separate sentences by . not ,

@@
 anecdotal evidence that the classifier work. To get a measure for the
-classifier's accuracy we retrain it in a cross-validation scheme, this
-yields the fraction of correct answers on three different folds:
+classifier's accuracy we retrain it in a cross-validation scheme.
+This yields the fraction of correct answers on three different folds:
 \begin{lstlisting}

### 2014-07-01 article.tex
typos

--- 2014-07-01 article.tex
@@
 and the corresponding algorithms slows down its adoption. Scikit-learn is
-a open-source software project that aims to make machine learning
+a open-source software project that aims at making machine learning
 accessible to all, whether it be in academia or in the industry. It
@@
 %
-Implementations mostly consisted of scattered piece of code to download
+Implementations mostly consisted of scattered pieces of code to download
 on researchers' web pages, with the notable exception of
 statistic-specific environments such as the R language \cite{Rmanual}, or
-the Weka java library \cite{hall2009weka}.
+the Weka Java library \cite{hall2009weka}.
 
-Scikit-learn aims to bridge the gap between machine-learning research and
+Scikit-learn aims at bridging the gap between machine-learning research and
 applications by providing a library, and not an environment, in a
@@
 Fig.\,\ref{fig:data} data about wages in the US, from \cite{berndt1991}:
-11 features, including wage, number of years of education and of work
-experience, on 534 individuals --samples in machine-learning
+11 \emph{features}, including wage, number of years of education and of work
+experience, on 534 individuals -- called \emph{samples} in machine-learning
 jargon.
@@
 cross\_val\_score} function measures the prediction score of an
-estimator. Cross-validation can also be used to tuned the meta-parameters
+estimator. Cross-validation can also be used to tune the meta-parameters
 of an estimator, such as the sparsity of a sparse model. For this,
@@
 Not every dataset comes as a set of numbers. For instance in natural
-language processing (NLP), observations are strings,
-which needs to be transformed into the data matrix expected by
-estimators. Often, features used to build decisions indicate the
-presence of certain words, or their frequencies.
+language processing (NLP), each observation is a text, \emph{i.e.}~a
+string, that needs to be transformed into the data matrix expected by
+estimators. Typically, features used to build decisions indicate the
+presence of certain words, and often their frequencies.
 In spam filtering, it can be very relevant to extract from a document that
-it contains the term \textsf{viagra}.
+it contains the term \textsf{viagra}, and possibly how many times.
 In other NLP problems, such as finding proper names,
-events such as ``next token is a Roman numeral'' may be relevant, as in
-``J.P. Kennedy III''.
-Reusing the {\tt transform} API, scikit-learn
-provides \emph{feature extraction} objects, that turn such
+events such as ``next token is the Roman numeral'' may be relevant, as in
+``J.P. Kennedy III''. A way to capture this information is to count bigrams,
+or $n$-grams. Reusing the {\tt transform} API, scikit-learn
+provides \emph{feature extraction} objects, that transform such
 unstructured data into a data matrix. In these situations, the
-$\textbf{X}$ matrix is a matrix of counting statistics.
-Since each word is typically present in only a few documents,
-the data matrix will be mostly zeros.
-SciPy offers a \emph{sparse matrix} package
-that stores the zeros in such matrices implicitly,
-and many scikit-learn estimators use that package to improve scalability.
+$\textbf{X}$ matrix is a
+matrix of counting statistics. It is full of zeros, in the sense that
+often a word is present only in a few documents. For better scalability,
+it can be encoded using SciPy's sparse matrices, that many scikit-learn
+estimators process efficiently.
 
@@
 learning from a dataset are all done in one programming language.
-This script first fetches a hand-labeled movie review dataset
+This script first fetches the hand-labeled movie review dataset from
 \cite{pang2004} and unpacks it to disk using standard Python libraries.
@@
 on word frequencies-- and logistic regression, and finally trains the
-model. Parameters for the feature extractor and the classifier
+model. Parameters for the feature extractor and the LogReg classifier
 were set manually.
@@
 particular to be computationally tractable. It is important to
-distinguish the case of very large number of samples, or very large
+distinguish between the case of very large number of samples and the case of large
 number of features. For a very large number of features, on-the-fly data
@@
 make better code and better design. From a developer's perspective, the
-project thus is managed with the goals of enabling newcomers to
+project is managed with the goals of enabling newcomers to
 contribute while keeping a high quality assurance. For these purposes, we
-rely heavily on Github (\url{http://github.com}) with an online
+rely heavily on Github --\url{http://github.com}-- with an online
 development workflow relying on distributed version control using git,
@@
 unit testing (up to 88\% of the lines of code are covered), run
-continuously on all contributions thanks to the travis platform
+continuously on all contributions thanks to the Travis platform
 (\url{http://travis-ci.org}).

### 2014-07-01 nlp_example.py

@@ -2,8 +2,8 @@ import urllib, tarfile  # Standard library imports
 from sklearn import linear_model, pipeline, datasets, feature_extraction
 
 urllib.urlretrieve('http://bit.ly/1juuXIr', 'tmp.tgz') # Download the data
-tarfile.open('tmp.tgz').extractall(path='.') # Extract it
-data = datasets.load_files('txt_sentoken') # And load it
+tarfile.open('tmp.tgz').extractall(path='.')           # Extract it
+data = datasets.load_files('txt_sentoken')             # And load it
 
 clf = pipeline.Pipeline([('step1', feature_extraction.text.TfidfVectorizer(min_df=2,
                                                  sublinear_tf=True, ngram_range=(1, 2))),

### 2014-07-01 article.tex

@@
 Not every dataset comes as a set of numbers. For instance in natural
-language processing (NLP), each observation is a text, \emph{i.e.}~a
-string, that needs to be transformed into the data matrix expected by
-estimators. Typically, features used to build decisions indicate the
-presence of certain words, and often their frequencies.
+language processing (NLP), observations are strings,
+which needs to be transformed into the data matrix expected by
+estimators. Often, features used to build decisions indicate the
+presence of certain words, or their frequencies.
 In spam filtering, it can be very relevant to extract from a document that
-it contains the term \textsf{viagra}, and possibly how many times.
+it contains the term \textsf{viagra}.
 In other NLP problems, such as finding proper names,
-events such as ``next token is the Roman numeral'' may be relevant, as in
-``J.P. Kennedy III''. A way to capture this information is to count bigrams,
-or $n$-grams. Reusing the {\tt transform} API, scikit-learn
-provides \emph{feature extraction} objects, that transform such
+events such as ``next token is a Roman numeral'' may be relevant, as in
+``J.P. Kennedy III''.
+Reusing the {\tt transform} API, scikit-learn
+provides \emph{feature extraction} objects, that turn such
 unstructured data into a data matrix. In these situations, the
-$\textbf{X}$ matrix is a
-matrix of counting statistics. It is full of zeros, in the sense that
-often a word is present only in a few documents. For better scalability,
-it can be encoded using SciPy's sparse matrices, that many scikit-learn
-estimators process efficiently.
+$\textbf{X}$ matrix is a matrix of counting statistics.
+Since each word is typically present in only a few documents,
+the data matrix will be mostly zeros.
+SciPy offers a \emph{sparse matrix} package
+that stores the zeros in such matrices implicitly,
+and many scikit-learn estimators use that package to improve scalability.
 
@@
 on word frequencies-- and logistic regression, and finally trains the
-model. Parameters for the feature extractor and the LogReg classifier
+model. Parameters for the feature extractor and the classifier
 were set manually.
@@
 anecdotal evidence that the classifier work. To get a measure for the
-classifier's accuracy we retraining it in a cross-validation scheme, this
-yields the fraction of correct answers on three different folds:
+classifier's accuracy we retrain it in a cross-validation scheme.
+This yields the fraction of correct answers on three different folds:
 \begin{lstlisting}

### 2014-07-01 article.tex
rewrite feature hashing part
The output vectors of the feature hashing algorithm aren't really
hash tables; they are more similar to Bloom filters.

@@
 is that the vectorization must be done in two passes over the documents,
-as the size of the output vector is not known before hand. Here, hashing,
-a trick similar to random projection, can be used to tackle both of these
-challenge and produce a streaming vectorizer, the {\tt HashingVectorizer}
-in scikit-learn. The idea is to use a hash table, storing hashes of the
-words, rather than the words themselves. The dimensionality of the hash
-table is known from the hash size.
+as the size of the output vectors is not known beforehand.
+Here, \emph{feature hashing} can be used
+to tackle both of these challenges and produce a streaming vectorizer,
+the {\tt HashingVectorizer} in scikit-learn.
+The idea is to run words through a hash function, rather than a lookup table,
+to map words to indices.
+Collisions may occur, but the hashing algorithm \cite{weinberger2009}
+is designed to mitigate their effect
+(and because words often have multiple meanings,
+``semantic'' collisions are inevitable in text processing tasks
+\cite{attenberg2009}).
+The dimensionality of the output vectors can be set by the user;
+the bigger, the less likely collisions will occur.
 

### 2014-07-01 article.tex
grammar + typo

@@
 Scikit-learn started circa 2007 as code from David Cournapeau and
-Matthieu Brucher's PhD work, but dwindled down till 2010, year at which the
+Matthieu Brucher's PhD work, but dwindled down till 2010, in which year the
 Parietal team from INRIA adopted the project, hiring a full time
-engineer. Basics API were defined and an efficient binding of LibSVM
+engineer. Basic APIs were defined and an efficient binding of LibSVM
 \cite{chang2011libsvm} gave the project a compelling advantage. In

### 2014-07-01 article.tex
random forests in lowercase
Uppercase variant is used as a trademark.

@@
 statistical analysis is extrapolation. The data is very noisy: for a
-given education and work experience wages vary widely. This variability
+given pair of education and work experience values, wages vary widely.
+This variability
 is probably explained by missing factors, such as sector of
@@
 On Fig.\,\ref{fig:random_forest}, we use a popular machine-learning
-algorithm, Random Forests, to predict the wage from the years of
+algorithm, random forests, to predict the wage from the years of
 education, or the years of work experience, either as separate features,

### 2014-07-01 article.tex
typos

@@
 inners of the algorithm, and it illustrates the corresponding
-extrapolation mechanism. The predictions fit the data very; maybe too
+extrapolation mechanism. The predictions fit the data well; maybe too
 well, as, on the prediction solely from the years of education (left of
@@
 positive, \emph{i.e.} negative, while the second one is positive. We have
-anecdotal evidence that the classifier work. To get a measure for the
+anecdotal evidence that the classifier works. To get a measure for the
 classifier's accuracy we retrain it in a cross-validation scheme.

### 2014-07-01 article.tex
grammar

@@
 Machine learning is about extracting rules from data, most often with the
-goal of making decision on new data \cite{elemstatlearn}. As a simple example, we show on
-Fig.\,\ref{fig:data} data about wages in the US, from \cite{berndt1991}:
+goal of making decision on new data \cite{elemstatlearn}. As a simple example, we show in
+Fig.~\ref{fig:data} data about wages in the US, from \cite{berndt1991}:
 11 \emph{features}, including wage, number of years of education and of work

### 2014-07-01 article.tex
in, not on, a figure

@@
 
-On Fig.\,\ref{fig:random_forest}, we use a popular machine-learning
+In Fig.~\ref{fig:random_forest}, we use a popular machine-learning
 algorithm, random forests, to predict the wage from the years of
@@
 
-On Fig.\,\ref{fig:linear_svm}, we use another popular machine-learning
+In Fig.~\ref{fig:linear_svm}, we use another popular machine-learning
 algorithm, linear Support Vector Machines (SVMs). Unlike random forests,
@@
 
-On Fig.\,\ref{fig:code} we give an example of fully functioning code to
+In Fig.~\ref{fig:code} we give an example of fully functioning code to
 do sentiment polarity classification for movie reviews: given a review,

### 2014-07-01 article.tex
trim use of --

@@
 most benefit from them: researchers --biologists, climate
-scientists, experimental physicists-- or developers --for web
+scientists, experimental physicists-- or developers of web
 services or domain-specific applications.
@@
 pipeline of a \textsf{tf--idf} feature extractor
-\cite{rennie2003tackling} --that turns strings into a sparse matrix based
-on word frequencies-- and logistic regression, and finally trains the
+\cite{rennie2003tackling} (that turns strings into a sparse matrix based
+on word frequencies) and logistic regression, and finally trains the
 model. Parameters for the feature extractor and the classifier
@@
 contributors, coming from different backgrounds, working in different
-institutions, with a variety of applications --more than 200 contributors
+institutions, with a variety of applications: more than 200 contributors in
 total, 15 active core contributors in the past year. Indeed, not only

### 2014-07-01 article.tex
capitalize NumPy, SciPy

@@
 have a {\tt fit} method, that accepts an $(n \times p)$ data matrix,
-$\mathbf{X}$ --numpy array or scipy sparse matrix--, where $n$ is the
+$\mathbf{X}$ --NumPy array or SciPy sparse matrix--, where $n$ is the
 number of samples, and $p$ the number of features--, and possibly a

### 2014-07-01 article.tex
minor stuff (kudos @Anne Schuth, UvA)

@@
 and the corresponding algorithms slows down its adoption. Scikit-learn is
-a open-source software project that aims at making machine learning
+an open-source software project that aims at making machine learning
 accessible to all, whether it be in academia or in the industry. It
@@
 Scikit-learn was born from the observation that most standard
-machine-learning algorithms were out of reach of the users that could
+machine learning algorithms were out of reach of the users that could
 most benefit from them: researchers --biologists, climate
@@
 %
-Implementations mostly consisted of scattered pieces of code to download
+Implementations of these algorithms
+mostly consisted of scattered pieces of code to download
 on researchers' web pages, with the notable exception of
@@
 funding from Google. Today, the project has grown vastly beyond
-INRIA into a worldwide open source effort.
+INRIA into a worldwide open source effort with more than 200 contributors.
 
@@
 11 \emph{features}, including wage, number of years of education and of work
-experience, on 534 individuals -- called \emph{samples} in machine-learning
-jargon.
+experience, describe 534 individuals,
+called \emph{samples} in machine-learning jargon.
 Code downloading the data to reproduce the figures can be found on
@@
 the figure), it is hard to believe in the bump at 5 years. This bump is
-probably a classic case of \emph{overfit}: the algorithm is learning its
+probably a classic case of \emph{overfitting}: the algorithm is learning its
 prediction from noise in the data. As a result, the prediction error on
@@
 following a linear law. The art of machine learning consists in choosing
-the right family of algorithms (or models) to describe the data well and
+the right family of algorithms/models to describe the data well and
 find the sweet spot between under and over fit. As the number of features
@@
 consumption by estimators.
+Scikit-learn comes with flexible feature extraction code to solve this problem.
 
@@
 %
-The data used to train a model can be used to measure the prediction
-error, as it would be enable to distinguish noise from signal, and thus
+The data used to train a model cannot be used to measure the prediction
+error, as it would be unable to distinguish noise from signal, and thus
 overly optimistic. The correct way is to leave out \emph{test} data,
@@
 in novelty detection, many are useful to transform data, as in dimension
-random. Scikit-learn estimators can have a {\tt transform} method, used
+reduction. Scikit-learn estimators can have a {\tt transform} method, used
 for this purpose. A {\tt Pipeline} estimator can then chain estimators to
@@
 In spam filtering, it can be very relevant to extract from a document that
-it contains the term \textsf{viagra}.
+it contains the term ``casino''.
 In other NLP problems, such as finding proper names,
 events such as ``next token is a Roman numeral'' may be relevant, as in
-``J.P. Kennedy III''.
+``J.P. Kennedy III.''
 Reusing the {\tt transform} API, scikit-learn
@@
 
-Here, we asked the estimator to predict if the reviews were positive for
-a batch of two reviews. The first review is predicted has being not
-positive, \emph{i.e.} negative, while the second one is positive. We have
+Here, we asked the estimator to predict whether the reviews were positive
+The first review is predicted as being negative (0),
+while the second is positive (1). We have
 anecdotal evidence that the classifier works. To get a measure for the

### 2014-07-01 article.tex
drop remark about "semantic collisions"

@@
 Collisions may occur, but the hashing algorithm \cite{weinberger2009}
-is designed to mitigate their effect
-(and because words often have multiple meanings,
-``semantic'' collisions are inevitable in text processing tasks
-\cite{attenberg2009}).
+is specially designed to mitigate their effect.
 The dimensionality of the output vectors can be set by the user;

### 2014-07-01 article.tex
restructure "Setting up models" paragraph

@@
 %
-Models in scikit-learn, called \emph{estimators}, are objects
-instantiated with all parameters that describe them, but with data. They
-have a {\tt fit} method, that accepts an $(n \times p)$ data matrix,
-$\mathbf{X}$ --NumPy array or SciPy sparse matrix--, where $n$ is the
-number of samples, and $p$ the number of features--, and possibly a
-quantity to predict, $\mathbf{y}$, also an array, 1D unless the
-prediction is \emph{multi-output}, with floats for a regression task, as
-in the wages prediction, or integers for classification, that is
-predicting for each sample to which class it belongs. The data seldom
-come in this form, and they must be adapted to be suitable for
-consumption by estimators.
-Scikit-learn comes with flexible feature extraction code to solve this problem.
+Learning algorithms in scikit-learn are embodied in \emph{estimators}:
+objects instantiated with parameters that control learning.
+Data is passed to the {\tt fit} method,
+which accepts an $(n \times p)$ data matrix $\mathbf{X}$,
+represented as a NumPy array or SciPy sparse matrix, where $n$ is the
+number of samples, and $p$ the number of features.
+When there is a quantity or class label to predict,
+a second argument $\mathbf{y}$ (usually a 1D array)
+contains the desired outcomes for the rows in $\mathbf{X}$.
+For a regression task such as wage prediction, these are real-valued,
+while in classification they are integers or strings that serve as labels
+for the class to which each sample belongs.
+Because data seldom arrives in this format,
+scikit-learn comes with flexible feature extraction code
+to make data suitable for consumption by estimators.
 

### 2014-07-01 article.tex
prediction "of some value of interest" (XXX or variable?)
cosmit

@@
 and numeric tools \cite{oliphant2007python,varoquaux2013scipy}, but also
-text processing tools, web servers...
+text processing tools, web servers, etc.
 %
@@
 %
-The goal of supervised models is prediction. The corresponding estimators
+The goal of supervised models is prediction of some value of interest.
+The corresponding estimators
 have a {\tt predict} method, that takes a data matrix $\mathbf{X}$ and

### 2014-07-01 article.tex
restructure big data section

@@
 %
-``Big data'' is the latest trend of data analytics: applying machine
+\emph{Big data} is a major trend in data analytics: applying machine
 learning to large datasets can extract information so far unexploited.
-These large datasets present new technical challenges, in
-particular to be computationally tractable. It is important to
-distinguish between the case of very large number of samples and the case of large
-number of features. For a very large number of features, on-the-fly data
-reductions, that exploit latent redundancy in the features, come in
-handy. In scikit-learn, one could for instance use
-random projections, that exploit the Johnson-Lindenstrauss lemma to give
-good approximations of Euclidean distances in the dataset. For
-a very large number of samples, some estimators expose
-\emph{on-line} algorithms: they can be fit incrementally, with successive
-calls to their {\tt partial\_fit} method, and converge in a small number
-of passes on the data. To alleviate one of the main practical challenges of
-big data, data-transfer latency, optimal data-fetching code can be used
-in parallel with the calls to {\tt partial\_fit}.
+While scikit-learn is most efficient in the ``medium data'' range,
+using shared memory multiprocessors rather than clusters,
+it has facilities for dealing with datasets that don't fit
+in any single machine.
 
-Dealing with a large corpus of text faces both challenges: many samples,
-and a very high dimensional space. In addition, the dimensionality
-tends to increase with sample size, as the number of different words used
-is greater in large corpuses. Beyond the mere size of the data matrix
-created by a standard vectorization approach, as above, another challenge
+Thinking in terms of the $(n \times p)$ data matrix,
+we can distinguish between the case of large numbers of samples $n$,
+and large numbers of features $p$.
+For a large number of features, scikit-learn's options include
+random projections, which reduce data dimensionality
+while preserving most of the geometric structure.
+For the ``large $n$'' case, some estimators expose a \texttt{partial\_fit}
+method that does \emph{online learning}.
+Instead of stashing all of the data in a single matrix,
+the user can feed the estimator slices of the matrix,
+and the algorithm finds a good approximation to its objective
+after a few passes over the dataset.
+
+% The citation is for "Heaps' law":
+% vocabulary size as a function of collection size.
+% Gael, feel free to remove it if you think it's too detailed.
+With large collections of text, one faces both challenges:
+the dimensionality $p$
+tends to increase with sample size $n$, as the number of different words used
+is greater in large collections \cite[88--89]{manning2009}.
+Beyond the mere size of the data matrix
+created by a standard vectorization approach, outlined earlier, another challenge
 is that the vectorization must be done in two passes over the documents,
-as the size of the output vectors is not known beforehand.
+as the size of the output vectors is not known up front.
 Here, \emph{feature hashing} can be used
@@
 the {\tt HashingVectorizer} in scikit-learn.
-The idea is to run words through a hash function, rather than a lookup table,
+Words are run through a (stateless) hash function, rather than a lookup table,
 to map words to indices.

### 2014-07-02 article.tex
Rephrase

@@
 method that does \emph{online learning}.
-Instead of stashing all of the data in a single matrix,
-the user can feed the estimator slices of the matrix,
+Instead of loading all of the data in a single huge matrix, the user
+can feed the estimator chunks of the data,
 and the algorithm finds a good approximation to its objective
@@
 no straightforward revenue model. Even more challenging is rewarding
-properly the long tail of small contributors. Finally, active
-contributors tend to be victim of the project's success, and be promoted
-to leading positions which leave them little time.
+properly the long tail of small contributors.
+%Finally, active
+%contributors tend to be victim of the project's success, and be promoted
+%to leading positions which leave them little time.
 

### 2014-07-02 article.tex
Rework intro to be less pronounced

@@
 %
-Implementations of these algorithms
+In the scientific world, implementations of these algorithms
 mostly consisted of scattered pieces of code to download
-on researchers' web pages, with the notable exception of
-statistic-specific environments such as the R language \cite{Rmanual}, or
-the Weka Java library \cite{hall2009weka}.
+on researchers' web pages. Unifying efforts could be found in
+statistic-specific environments and libraries such as the R
+language \cite{Rmanual}, the Weka Java library \cite{hall2009weka}, or
+the Shogun C++ library \cite{sonnenburg2010}, but we felt that these were
+too technical to reach our target users, not versed in statistics.
 
@@
 
+\begin{figure*}[b]
+% To have the float on the bottom, we need this declared on the page
+% before where it is going to be printed
+\lstinputlisting[lastline=12]{nlp_example.py}
+\caption{Simple text-processing code to download movie reviews from
+Internet and learn positive versus negative ratings.\label{fig:code}}
+\end{figure*}
+
 \paragraph{Setting up models and feeding them data.}
@@
 returns a predicted $\mathbf{y}$.
-
-\begin{figure*}[b]
-% To have the float on the bottom, we need this declared on the page
-% before where it is going to be printed
-\lstinputlisting[lastline=12]{nlp_example.py}
-\caption{Simple text-processing code to download movie reviews from
-Internet and learn positive versus negative ratings.\label{fig:code}}
-\end{figure*}
 

### 2014-07-02 article.tex
model evaluation

@@
 %
-The data used to train a model cannot be used to measure the prediction
-error, as it would be unable to distinguish noise from signal, and thus
-overly optimistic. The correct way is to leave out \emph{test} data,
-untouched during training. Often a \emph{cross-validation} scheme is
+The performance of a model measures its ability to correctly
+predict new data. Because of overfitting, the \emph{train} data used to build a
+model should in general not be used to also evaluate its performance, as it may often
+result in overly optimistic estimates of its true prediction error.
+Accordingly, the correct way to obtain an unbiased estimate is to leave out \emph{test} data,
+untouched during training and used for model evaluation only.
+Often a \emph{cross-validation} scheme is
 used, where the data is repeatedly split into \emph{train} and
@@
 contribute while keeping a high quality assurance. For these purposes, we
-rely heavily on Github (\url{http://github.com}) with an online
+rely heavily on GitHub (\url{http://github.com}) with an online
 development workflow relying on distributed version control using git,

### 2014-07-02 article.tex
wording

@@
 The performance of a model measures its ability to correctly
-predict new data. Because of overfitting, the \emph{train} data used to build a
-model should in general not be used to also evaluate its performance, as it may often
-result in overly optimistic estimates of its true prediction error.
+predict new data. Because of overfitting, the \emph{train} data used for building
+should in general not be used to also evaluate model performance, as it may often
+result in overly optimistic estimates of the true prediction error of the model.
 Accordingly, the correct way to obtain an unbiased estimate is to leave out \emph{test} data,
 untouched during training and used for model evaluation only.
-Often a \emph{cross-validation} scheme is
+Alternatively, a \emph{cross-validation} scheme may also be
 used, where the data is repeatedly split into \emph{train} and

### 2014-07-02 article.tex
wording

@@
 result in overly optimistic estimates of the true prediction error of the model.
-Accordingly, the correct way to obtain an unbiased estimate is to leave out \emph{test} data,
+Instead, the correct way to obtain an unbiased estimate is to leave out \emph{test} data,
 untouched during training and used for model evaluation only.

### 2014-07-03 article.tex
better defense of Python as implementation language
Speaking for myself at least, the problem with R and Weka is not
so much the statistics but the programming languages: R is a
poorly designed language, Weka uses two different interfaces
(GUI and Java).

I dropped Shogun because I'm not sure if it has the same audience
as the others. It seems to focus on kernel learning and very large
datasets.

@@
 statistic-specific environments and libraries such as the R
-language \cite{Rmanual}, the Weka Java library \cite{hall2009weka}, or
-the Shogun C++ library \cite{sonnenburg2010}, but we felt that these were
-too technical to reach our target users, not versed in statistics.
+language \cite{Rmanual} or the Weka Java toolkit \cite{hall2009weka}.
+Of these, R is a custom \emph{language},
+while Java is a library with a custom GUI.
+The choice of Python ties machine learning into a general-purpose language
+that allows interactive programming \cite{perez2007ipython},
+with a single interface for
+interactive work, scripting and full-blown applications.
 

### 2014-07-04 article.tex
Wording

--- 2014-07-02 article.tex
@@
 In the scientific world, implementations of these algorithms
-mostly consisted of scattered pieces of code to download
+mostly consisted of scattered pieces of code %to download
 on researchers' web pages. Unifying efforts could be found in
 statistic-specific environments and libraries such as the R
-language \cite{Rmanual}, the Weka Java library \cite{hall2009weka}, or
-the Shogun C++ library \cite{sonnenburg2010}, but we felt that these were
-too technical to reach our target users, not versed in statistics.
+language \cite{Rmanual}, the Weka Java toolkit \cite{hall2009weka}, or the
+Shogun C++ library \cite{sonnenburg2010}.
+Of these, R is a custom \emph{language},
+while Java is a library with a custom GUI, and Shogun is a fairly
+technical C++ code base.
+In scikit-learn, we chose Python to reach many users, technical or not, tying machine learning
+into a general-purpose language good for simple
+ interactive programming \cite{perez2007ipython}, sophisticated
+scripting, as well as full-blown applications.
 
@@
 matplotlib \cite{hunter2007matplotlib} --for scientific plotting-- and
-Cython, to generate and call compiled code in a Python-like syntax
+Cython, to generate compiled code in a Python-like syntax
 \cite{behnel2011cython}.
@@
 
-\begin{figure}[b]
+\begin{figure}[t]
     %\hspace*{-.015\linewidth}%
@@
     education; data from \cite{berndt1991}\label{fig:data}.}
+\end{figure}
+
+\begin{figure}[b]
+    %\hspace*{-.015\linewidth}%
+    \includegraphics[width=1.05\linewidth]{wage_data_random_forest}%
+
+    \caption{Wage prediction from years of work experience and education,
+    using random forests.\label{fig:random_forest}}
 \end{figure}
@@
 training data.
-
-\begin{figure}[b]
-    %\hspace*{-.015\linewidth}%
-    \includegraphics[width=1.05\linewidth]{wage_data_random_forest}%
-
-    \caption{Wage prediction from years of work experience and education,
-    using random forests.\label{fig:random_forest}}
-\end{figure}
 
@@
 The performance of a model measures its ability to correctly
-predict new data. Because of overfitting, the \emph{train} data used for building
+predict new data. Because of overfitting, the \emph{train} data used for
+learning
 should in general not be used to also evaluate model performance, as it may often
@@
 used, where the data is repeatedly split into \emph{train} and
-\emph{test} subsets. Scikit-learn provides a complete framework for
-cross-validation. Specific iterators are used to describe the subsets and
-several functions and objects accept as arguments these iterators to
+\emph{test} subsets. Scikit-learn provides integrated support for
+cross-validation. Specific iterators can specify the train and test
+subsets. Several functions and objects accept these iterators as arguments to
 perform cross-validation internally. For instance the {\tt
@@
 the data matrix will be mostly zeros.
-SciPy offers a \emph{sparse matrix} package
+SciPy offers a \emph{sparse matrix} object
 that stores the zeros in such matrices implicitly,
-and many scikit-learn estimators use that package to improve scalability.
+and many scikit-learn estimators can use this object to improve scalability.
 
@@
 industry. Indeed, getting due credit for grass-root, free, software that
-"just works" is difficult. None of the grant proposals submitted to fund
+``just works'' is difficult. None of the grant proposals submitted to fund
 development by core contributors have so far been accepted, and there is
@@
 methods most useful for prediction tasks. We favor time-tested
-algorithms and put great care in the choice of default parameters. Our
-secret ambition is that the documentation, that we strive to make
-didactic and pragmatic, can provide an executable alternative to
+algorithms and put great care in the choice of default parameters. We put
+a lot of effort in making the documentation didactic and
+pragmatic. Our
+secret ambition is that it can provide an executable alternative to
 textbooks, with all the figures generated by code examples.

### 2014-07-04 article.tex
wording

@@
 The code demonstrates the use of pipelines, and also shows how machine
-learning ties into unrelated Python's modules: downloading, unpacking and
+learning ties into other Python modules: downloading, unpacking and
 learning from a dataset are all done in one programming language.

### 2014-07-04 article.tex
wording

@@
 
-Here, we asked the estimator to predict whether the reviews were positive
+Here, we asked the estimator to predict whether the reviews were
+positive.
 The first review is predicted as being negative (0),

### 2014-07-06 article.tex
Minor formulation

@@
 classifier's accuracy we retrain it in a cross-validation scheme.
-This yields the fraction of correct answers on three different folds:
+This scheme yields the fraction of correct answers on three different folds:
 \begin{lstlisting}
@@
 
-The open-source, community-driven model of scikit-learn has fantastically
+The open-source community-driven model of scikit-learn has fantastically
 payed off. However, we find that it does not fit in the standard

### 2014-07-06 article.tex
Tentative author list and affiliations

@@
 
-\author{The authors}
+%\author{G. Varoquaux \and L. Buitinck \and G. Louppe \and O. Grisel \and F.
+%Pedregosa \and A. Mueller}
+
+\author[1]{G. Varoquaux\thanks{gael.varoquaux@inria.fr}}
+\author[2]{L. Buitinck}
+\author[3]{G. Louppe}
+\author[1]{O. Grisel}
+\author[1]{F. Pedregosa}
+\author[4]{A. Mueller}
+\affil[1]{Parietal, INRIA, Bat 145 CEA Saclay, 91191 Gif-sur-Yvette,
+France}
+\affil[2]{ILPS, University of Amsterdam, Science Park 904, Amsterdam, The
+Netherlands}
+\affil[3]{Systems and Modeling Research Unit, University of Liège
+Montefiore Institute (B28), Liège, Belgium}
+\affil[4]{Amazon Development Center Germany, Berlin, Germany}
 

### 2014-11-04 article.tex
Try to address some of the comments of the editor

@@
 
+Machine learning is a very active field of research, with promising
+trends such as deep learning, harnessing GPUs or distributed computing
+for extra computational power, or extending to new platforms, such as
+Android or iOS-based embedded devises. However, as scikit-learn purpose
+is to provided tried-and-true, mostly turn-key, predictive tools, we have
+to walk the fine line between supporting powerful techniques and
+utilizing our limited project cycles on proven technologies. For
+instance, deep neural networks are an emerging tool with a huge
+potential, however, they are very much a technology in flux and currently
+are far from being usable with a significant expertise. Similarly, GPUs
+give a lot of user-level problems, as there is a lack of binary interface
+to address them uniformly and robustly over all hardwares. Finally, we
+support new platforms, such as Android, as long as they come with a good
+Python implementation with a standard build chain. Indeed, our basic
+execution environment is the scientific Python stack, that is both very
+powerful and high-level. The vision of the scikit-learn project is to
+bring advances in machine learning to users that that may lack the full
+expertise. As a consequence, we focus on seasoned approaches, rather than
+implementing ongoing research. We believe that code is an intrinsic part
+of the scholarship outcome of scientific research in computational
+science \cite{buckheit1995wavelab}, and scikit-learn strives to be akin
+to a reference textbook in this research software landscape.
+
 %We have found that implementing, even a standard algorithm, really well,

### 2014-11-06 article.tex
FIX: make example copy/pastable

@@
 \usepackage{magazine-style}
+\usepackage{textcomp}
+
+\makeatletter
+\newcommand*\verbfont{%
+    \normalfont\ttfamily\hyphenchar\font\m@ne\@noligs}
+\makeatother
 
@@
 	extendedchars=true,
+    columns=fullflexible,
+    upquote=true,
 	xleftmargin = 0pt,
@@
         belowskip = 0.6ex,
-	basicstyle=\small\sffamily,
+	basicstyle=\verbfont,
 	escapebegin={\color{green}},
@@
 	%identifierstyle=\sffamily\small,
-	commentstyle=\color{green!40!black},
-	stringstyle=\rmfamily\color{red!60!black},
+	commentstyle=\color{green!40!black} \verbfont,
+	stringstyle=\color{red!60!black} \verbfont,
 	showstringspaces=false,

### 2014-11-06 article.tex
how to plug a different classifier into the NLP example

--- 2014-11-04 article.tex
@@
 
+We see that the expected accuracy is about 88\%.
+Experimenting with different classification models is easy:
+e.g., replacing \texttt{LogisticRegression} with \texttt{svm.LinearSVC}
+gives a linear support vector machine.
+Changing the simple bag-of-words representation to a \textit{bigram}
+(pairs of adjacent words) one is a matter of giving one additional argument
+to the tf--idf vectorizer, which reads \texttt{n\_gram\_range=(2, 2)}.
+(Similarly, the value \texttt{(2,3)} would extract bigrams and trigrams, etc.)
+
 \paragraph{What about big data?}

### 2014-11-06 article.tex
actually n-grams aren't that interesting
nurturing section: stress code review over version control

--- 2014-11-04 article.tex
@@
 
+We see that the expected accuracy is about 88\%.
+Experimenting with different classification models is easy:
+e.g., replacing \texttt{LogisticRegression} with \texttt{svm.LinearSVC}
+gives a linear support vector machine.
+
 \paragraph{What about big data?}
@@
 total, 15 active core contributors in the past year. Indeed, not only
-is there strength in numbers, but also the convergence of points of views
-make better code and better design. From a developer's perspective, the
-project is managed with the goals of enabling newcomers to
-contribute while keeping a high quality assurance. For these purposes, we
-rely heavily on GitHub (\url{http://github.com}) with an online
-development workflow relying on distributed version control using git,
-and intensive code review. The code review process is a great asset to
+is there strength in numbers, but the convergence of points of views
+make for better code and better design. From a developer's perspective, the
+project is managed with the goal of enabling newcomers to contribute,
+while keeping quality high.
+
+Our development workflow relies on intensive code review,
+facilitated by the online version control environment GitHub (\url{http://github.com})
+The code review process is a great asset to
 raising the quality of the code, from all points of view: algorithmic,

### 2014-11-06 article.tex
Fix typo (Java -> Weka)

@@
 France}
-\affil[2]{ILPS, University of Amsterdam, Science Park 904, Amsterdam, The
+\affil[2]{ISLA, University of Amsterdam, Science Park 904, Amsterdam, The
 Netherlands}
@@
 Of these, R is a custom \emph{language},
-while Java is a library with a custom GUI, and Shogun is a fairly
+while Weka is a Java library with a custom GUI, and Shogun is a fairly
 technical C++ code base.
@@
 Scikit-learn started circa 2007 as code from David Cournapeau and
-Matthieu Brucher's PhD work, but dwindled down till 2010, in which year the
-Parietal team from INRIA adopted the project, hiring a full time
+Matthieu Brucher's PhD work, but dwindled down until 2010, when the
+Parietal team at INRIA adopted the project, hiring a full-time
 engineer. Basic APIs were defined and an efficient binding of LibSVM
@@
 \lstinputlisting[lastline=12]{nlp_example.py}
-\caption{Simple text-processing code to download movie reviews from
-Internet and learn positive versus negative ratings.\label{fig:code}}
+\caption{Simple text-processing code to download movie reviews
+and learn positive versus negative ratings.\label{fig:code}}
 \end{figure*}
@@
 numerical robustness, homogeneity of APIs, ease of use, documentation.
-Another central aspect for quality in the project is the heavy use of
-unit testing (up to 88\% of the lines of code are covered), run
-continuously on all contributions thanks to the Travis platform
+Another central tool for maintaining quality in the project is the use of
+unit testing (some 88\% of the lines of code are covered).
+All would-be code contributions are automatically tested using Travis
 (\url{http://travis-ci.org}).
@@
 should offer a wide variety of learning algorithms. However, each new
-feature, each new line of code, comes with a significant maintenance cost.
+feature, each new line of code, comes with a maintenance cost.
 Uncontrolled growth would lead the project to pick up weight until it
@@
 secret ambition is that it can provide an executable alternative to
-textbooks, with all the figures generated by code examples.
+textbooks, with all the figures are generated by code examples.
 
@@
 for extra computational power, or extending to new platforms, such as
-Android or iOS-based embedded devises. However, as scikit-learn purpose
+Android or iOS-based embedded devises. However, as scikit-learn's purpose
 is to provided tried-and-true, mostly turn-key, predictive tools, we have
@@
 instance, deep neural networks are an emerging tool with a huge
-potential, however, they are very much a technology in flux and currently
-are far from being usable with a significant expertise. Similarly, GPUs
-give a lot of user-level problems, as there is a lack of binary interface
-to address them uniformly and robustly over all hardwares. Finally, we
-support new platforms, such as Android, as long as they come with a good
-Python implementation with a standard build chain. Indeed, our basic
+potential, but are still in flux and hard to use without significant expertise.
+
+Similarly, GPGPU technology is promising, but still poses user-level problems.
+While OpenCL largely hides the incompatibility of GPUs and other compute devices
+at the source code level \cite{stone2010opencl}, it does not guarantee performance portability
+and requires the installation of device-specific drivers.
+PyOpenCL \cite{klockner2012pycuda} may alleviate the former problem and communicates
+with NumPy, making it an obvious starting point for experimentation
+in the direction of GPU-enabled scikit-learn. Theano \cite{bergstra2010theano} has
+shown the feasibility of programming GPU-powered neural nets in Python.
+
+Support for mobile platforms such as Android is determined by those platforms'
+support for Python and the standard C/C++ build chain. Indeed, our basic
 execution environment is the scientific Python stack, that is both very

### 2014-11-06 article.tex
be a bit more positive about GPUs

@@
 instance, deep neural networks are an emerging tool with a huge
-potential, however, they are very much a technology in flux and currently
-are far from being usable with a significant expertise. Similarly, GPUs
-give a lot of user-level problems, as there is a lack of binary interface
-to address them uniformly and robustly over all hardwares. Finally, we
-support new platforms, such as Android, as long as they come with a good
-Python implementation with a standard build chain. Indeed, our basic
+potential, but are still in flux and hard to use without significant expertise.
+
+Similarly, GPGPU technology is promising, but still poses user-level problems.
+While OpenCL largely hides the incompatibility of GPUs and other compute devices
+at the source code level \cite{stone2010opencl}, it does not guarantee performance portability
+and requires the installation of device-specific drivers.
+PyOpenCL \cite{klockner2012pycuda} may alleviate the former problem and communicates
+with NumPy, making it the obvious starting point for experimentation
+in the direction of GPU-enabled scikit-learn. Theano \cite{bergstra2010theano} has
+shown the feasibility of programming GPU-powered neural nets in Python.
+
+Support for mobile platforms such as Android is determined by those platforms'
+support for Python and the standard C/C++ build chain. Indeed, our basic
 execution environment is the scientific Python stack, that is both very

### 2014-11-06 nlp_example.py
example: use make_pipeline for brevity, copyedit comments

@@ -1,13 +1,13 @@
 import urllib, tarfile  # Standard library imports
 from sklearn import linear_model, pipeline, datasets, feature_extraction
 
-urllib.urlretrieve('http://bit.ly/1juuXIr', 'tmp.tgz') # Download the data
-tarfile.open('tmp.tgz').extractall(path='.')           # Extract it
-data = datasets.load_files('txt_sentoken')             # And load it
+urllib.urlretrieve('http://bit.ly/1juuXIr', 'tmp.tgz') # Download the data,
+tarfile.open('tmp.tgz').extractall(path='.')           # extract it,
+data = datasets.load_files('txt_sentoken')             # and load it.
 
-clf = pipeline.Pipeline([('step1', feature_extraction.text.TfidfVectorizer(min_df=2,
-                                                 sublinear_tf=True, ngram_range=(1, 2))),
-                         ('step2', linear_model.LogisticRegression(C=5000))])
+clf = pipeline.make_pipeline(feature_extraction.text.TfidfVectorizer(min_df=2,
+                                                 sublinear_tf=True, ngram_range=(1, 2)),
+                             linear_model.LogisticRegression(C=5))
 clf.fit(data.data, data.target)
 
 # Demo

### 2014-11-06 nlp_example.py
oops, mistakenly changed the SVM parameters

@@ -7,7 +7,7 @@ data = datasets.load_files('txt_sentoken')             # and load it.
 
 clf = pipeline.make_pipeline(feature_extraction.text.TfidfVectorizer(min_df=2,
                                                  sublinear_tf=True, ngram_range=(1, 2)),
-                             linear_model.LogisticRegression(C=5))
+                             linear_model.LogisticRegression(C=5000))
 clf.fit(data.data, data.target)
 
 # Demo

### 2014-11-06 nlp_example.py
Revert "example: use make_pipeline for brevity, copyedit comments"
This reverts commit ca57d8e36d5918170256a2e48267fabf43da5f84.

Conflicts:
	nlp_example.py

@@ -5,9 +5,15 @@ urllib.urlretrieve('http://bit.ly/1juuXIr', 'tmp.tgz') # Download the data,
 tarfile.open('tmp.tgz').extractall(path='.')           # extract it,
 data = datasets.load_files('txt_sentoken')             # and load it.
 
+<<<<<<< HEAD
 clf = pipeline.make_pipeline(feature_extraction.text.TfidfVectorizer(min_df=2,
                                                  sublinear_tf=True, ngram_range=(1, 2)),
                              linear_model.LogisticRegression(C=5000))
+=======
+clf = pipeline.Pipeline([('step1', feature_extraction.text.TfidfVectorizer(min_df=2,
+                                                 sublinear_tf=True, ngram_range=(1, 2))),
+                         ('step2', linear_model.LogisticRegression(C=5000))])
+>>>>>>> parent of ca57d8e... example: use make_pipeline for brevity, copyedit comments
 clf.fit(data.data, data.target)
 
 # Demo

### 2014-11-06 article.tex
ILPS -> ISLA (Intelligent Systems Lab Amsterdam)
grammar fixes
GPGPU, again; address Gaël's comments

--- 2014-11-06 article.tex
@@
 potential, but are still in flux and hard to use without significant expertise.
-
-Similarly, GPGPU technology is promising, but still poses user-level problems.
-While OpenCL largely hides the incompatibility of GPUs and other compute devices
-at the source code level \cite{stone2010opencl}, it does not guarantee performance portability
-and requires the installation of device-specific drivers.
-PyOpenCL \cite{klockner2012pycuda} may alleviate the former problem and communicates
-with NumPy, making it an obvious starting point for experimentation
-in the direction of GPU-enabled scikit-learn. Theano \cite{bergstra2010theano} has
-shown the feasibility of programming GPU-powered neural nets in Python.
+Similarly, GPGPU technology is promising, but not ready for non-expert use.
+The HPC community and industry have largely achieved source-level compatibility with OpenCL
+\cite{stone2010opencl},
+but still need to solve the problems of performance portability and cross-platform driver support.
 

### 2014-11-06 nlp_example.py
botched merge (sorry!)

@@ -5,15 +5,9 @@ urllib.urlretrieve('http://bit.ly/1juuXIr', 'tmp.tgz') # Download the data,
 tarfile.open('tmp.tgz').extractall(path='.')           # extract it,
 data = datasets.load_files('txt_sentoken')             # and load it.
 
-<<<<<<< HEAD
-clf = pipeline.make_pipeline(feature_extraction.text.TfidfVectorizer(min_df=2,
-                                                 sublinear_tf=True, ngram_range=(1, 2)),
-                             linear_model.LogisticRegression(C=5000))
-=======
 clf = pipeline.Pipeline([('step1', feature_extraction.text.TfidfVectorizer(min_df=2,
                                                  sublinear_tf=True, ngram_range=(1, 2))),
                          ('step2', linear_model.LogisticRegression(C=5000))])
->>>>>>> parent of ca57d8e... example: use make_pipeline for brevity, copyedit comments
 clf.fit(data.data, data.target)
 
 # Demo

### 2014-11-06 article.tex
Slight compresson of the GP-GPU part

@@
 \cite{stone2010opencl},
-but still need to solve the problems of performance portability and cross-platform driver support.
-
+but performance portability and cross-platform driver support remain
+challenging.
+%
 Support for mobile platforms such as Android is determined by those platforms'

### 2014-11-06 article.tex
cosmetics (mostly re-aligning in the example script)

@@
 Netherlands}
-\affil[3]{Systems and Modeling Research Unit, University of Liège
-Montefiore Institute (B28), Liège, Belgium}
+\affil[3]{Montefiore Institute, University of Liège, Liège, Belgium}
 \affil[4]{Amazon Development Center Germany, Berlin, Germany}
@@
 applications by providing a library, and not an environment, in a
-general-purpose language, relying on domain-agnostic data structures
+general-purpose programming language, relying on domain-agnostic data structures
 \cite{pedregosa2011}. Emphasis is put on quality and ease of use, which
@@
 %
-The goal of supervised models is prediction of some value of interest.
+The goal of supervised models is the prediction of some value of interest.
 The corresponding estimators
@@
 \emph{test} subsets. Scikit-learn provides integrated support for
-cross-validation. Specific iterators can specify the train and test
+cross-validation, as specific iterators defining the train and test
 subsets. Several functions and objects accept these iterators as arguments to
@@
 do sentiment polarity classification for movie reviews: given a review,
-it outputs whether the review is positive about the movie it concerns.
+it outputs whether that review is positive about the movie it concerns.
 The code demonstrates the use of pipelines, and also shows how machine
@@
 were set manually.
+
+\begin{figure*}[b]
+% To have the float on the bottom, we need this declared on the page
+% before where it is going to be printed
+\lstinputlisting{online_learning_example.py}
+\caption{On-line learning example: the main loop iterates through batches
+of 1000 documents, vectorizing them on the fly and updating the
+classifier. Running the example shows that the error on the test set
+progressively decreases as data is accumulated.\label{fig:online}}
+\end{figure*}
 
@@
 the bigger, the less likely collisions will occur.
+%
+A full example doing online learning on a large corpus of text is given
+in Fig.~\ref{fig:online}.
 
@@
 secret ambition is that it can provide an executable alternative to
-textbooks, with all the figures are generated by code examples.
+textbooks, with all the figures generated by code examples.
 
@@
 for extra computational power, or extending to new platforms, such as
-Android or iOS-based embedded devises. However, as scikit-learn's purpose
-is to provided tried-and-true, mostly turn-key, predictive tools, we have
+Android or iOS-based embedded devices. However, as scikit-learn's purpose
+is to provide tried-and-true, mostly turn-key, predictive tools, we have
 to walk the fine line between supporting powerful techniques and

### 2014-11-06 nlp_example.py

@@ -1,13 +1,14 @@
-import urllib, tarfile  # Standard library imports
+import urllib, tarfile  # From the Python standard library
 from sklearn import linear_model, pipeline, datasets, feature_extraction
 
-urllib.urlretrieve('http://bit.ly/1juuXIr', 'tmp.tgz') # Download the data,
-tarfile.open('tmp.tgz').extractall(path='.')           # extract it,
-data = datasets.load_files('txt_sentoken')             # and load it.
+# Download, unpack and load the dataset.
+urllib.urlretrieve('http://bit.ly/1juuXIr', 'tmp.tgz')
+tarfile.open('tmp.tgz').extractall(path='.')
+data = datasets.load_files('txt_sentoken')
 
 clf = pipeline.Pipeline([('step1', feature_extraction.text.TfidfVectorizer(min_df=2,
-                                                 sublinear_tf=True, ngram_range=(1, 2))),
-                         ('step2', linear_model.LogisticRegression(C=5000))])
+                                                        sublinear_tf=True, ngram_range=(1, 2))),
+                            ('step2', linear_model.LogisticRegression(C=5000))])
 clf.fit(data.data, data.target)
 
 # Demo

### 2014-11-06 out_of_core_classification.py
WIP: on-line learning example

@@ -1,0 +1,31 @@
+from __future__ import print_function
+
+from sklearn.datasets import fetch_20newsgroups
+from sklearn.feature_extraction.text import HashingVectorizer
+
+# Create the hasher
+hasher = HashingVectorizer(decode_error='ignore', n_features=2 ** 16)
+
+dataset = fetch_20newsgroups(shuffle=True, random_state=1,
+                             remove=('headers', 'footers', 'quotes'))
+
+X_test = hasher.fit_transform(dataset.data[:1000])
+y_test = dataset.target[:1000]
+
+from sklearn.linear_model import PassiveAggressiveClassifier
+# SVM classifier that learns online using SGD.
+classifier = PassiveAggressiveClassifier(C=.01)
+
+from sklearn.linear_model import SGDClassifier
+# SVM classifier that learns online using SGD.
+classifier = SGDClassifier(learning_rate='invscaling', eta0=1.)
+
+# Main loop : iterate on mini-batchs of 1000 samples
+i = 1000
+while i < len(dataset.data) - 2000:
+    y_train = dataset.target[i:i + 1000]
+    X_train = hasher.fit_transform(dataset.data[i:i + 1000])
+    # update estimator with examples in the current mini-batch
+    classifier.partial_fit(X_train, y_train,
+                           classes=range(len(dataset.target_names)))
+    print(classifier.score(X_test, y_test))

### 2014-11-06 out_of_core_classification.py
Fix stupid bug (and change learner)

@@ -14,17 +14,17 @@ y_test = dataset.target[:1000]
 
 from sklearn.linear_model import PassiveAggressiveClassifier
 # SVM classifier that learns online using SGD.
-classifier = PassiveAggressiveClassifier(C=.01)
+classifier = PassiveAggressiveClassifier(C=.1)
 
-from sklearn.linear_model import SGDClassifier
-# SVM classifier that learns online using SGD.
-classifier = SGDClassifier(learning_rate='invscaling', eta0=1.)
+#from sklearn.linear_model import SGDClassifier
+## SVM classifier that learns online using SGD.
+#classifier = SGDClassifier(learning_rate='invscaling', eta0=1.)
 
 # Main loop : iterate on mini-batchs of 1000 samples
-i = 1000
-while i < len(dataset.data) - 2000:
-    y_train = dataset.target[i:i + 1000]
-    X_train = hasher.fit_transform(dataset.data[i:i + 1000])
+i = 1
+while i * 1000 < len(dataset.data) - 2000:
+    y_train = dataset.target[i * 1000:(i + 1) * 1000]
+    X_train = hasher.fit_transform(dataset.data[i * 1000:(i + 1) * 1000])
     # update estimator with examples in the current mini-batch
     classifier.partial_fit(X_train, y_train,
                            classes=range(len(dataset.target_names)))

### 2014-11-06 out_of_core_classification.py
More useless bug (this clearly needs more work)

@@ -29,3 +29,4 @@ while i * 1000 < len(dataset.data) - 2000:
     classifier.partial_fit(X_train, y_train,
                            classes=range(len(dataset.target_names)))
     print(classifier.score(X_test, y_test))
+    i += 1

### 2014-11-07 out_of_core_classification.py
Somewhat happy with the on-line learning example

@@ -1,5 +1,3 @@
-from __future__ import print_function
-
 from sklearn.datasets import fetch_20newsgroups
 from sklearn.feature_extraction.text import HashingVectorizer
 
@@ -12,21 +10,21 @@ dataset = fetch_20newsgroups(shuffle=True, random_state=1,
 X_test = hasher.fit_transform(dataset.data[:1000])
 y_test = dataset.target[:1000]
 
-from sklearn.linear_model import PassiveAggressiveClassifier
+from sklearn.linear_model import SGDClassifier
 # SVM classifier that learns online using SGD.
-classifier = PassiveAggressiveClassifier(C=.1)
+classifier = SGDClassifier()
 
-#from sklearn.linear_model import SGDClassifier
-## SVM classifier that learns online using SGD.
-#classifier = SGDClassifier(learning_rate='invscaling', eta0=1.)
+# Artificially make long train data
+train_data = 15 * dataset.data[1000:]
+train_target = 15 * list(dataset.target[1000:])
 
 # Main loop : iterate on mini-batchs of 1000 samples
-i = 1
-while i * 1000 < len(dataset.data) - 2000:
-    y_train = dataset.target[i * 1000:(i + 1) * 1000]
-    X_train = hasher.fit_transform(dataset.data[i * 1000:(i + 1) * 1000])
+i = 0
+while i < len(train_data) - 1000:
+    y_train = train_target[i:i + 1000]
+    X_train = hasher.fit_transform(train_data[i:i + 1000])
     # update estimator with examples in the current mini-batch
     classifier.partial_fit(X_train, y_train,
                            classes=range(len(dataset.target_names)))
-    print(classifier.score(X_test, y_test))
-    i += 1
+    print classifier.score(X_test, y_test)
+    i += 1000

### 2014-11-07 out_of_core_classification.py
Integrate online learning example

@@ -1,30 +1,22 @@
 from sklearn.datasets import fetch_20newsgroups
 from sklearn.feature_extraction.text import HashingVectorizer
 
-# Create the hasher
-hasher = HashingVectorizer(decode_error='ignore', n_features=2 ** 16)
+vectorizer = HashingVectorizer(decode_error='ignore', n_features=2 ** 16) # A vectorizer
 
-dataset = fetch_20newsgroups(shuffle=True, random_state=1,
-                             remove=('headers', 'footers', 'quotes'))
-
-X_test = hasher.fit_transform(dataset.data[:1000])
+dataset = fetch_20newsgroups(shuffle=True, random_state=1, remove=('headers', 'footers'))
+X_test = vectorizer.fit_transform(dataset.data[:1000])
 y_test = dataset.target[:1000]
 
 from sklearn.linear_model import SGDClassifier
-# SVM classifier that learns online using SGD.
-classifier = SGDClassifier()
+classifier = SGDClassifier()     # SVM classifier that learns online using SGD.
 
 # Artificially make long train data
-train_data = 15 * dataset.data[1000:]
-train_target = 15 * list(dataset.target[1000:])
+train_data = 10 * dataset.data[1000:]
+train_target = 10 * list(dataset.target[1000:])
 
-# Main loop : iterate on mini-batchs of 1000 samples
-i = 0
-while i < len(train_data) - 1000:
+for i in range(0, len(train_data), 1000):    # Iterate on mini-batchs of 1000 samples
     y_train = train_target[i:i + 1000]
-    X_train = hasher.fit_transform(train_data[i:i + 1000])
+    X_train = vectorizer.fit_transform(train_data[i:i + 1000])
     # update estimator with examples in the current mini-batch
-    classifier.partial_fit(X_train, y_train,
-                           classes=range(len(dataset.target_names)))
+    classifier.partial_fit(X_train, y_train, classes=range(len(dataset.target_names)))
     print classifier.score(X_test, y_test)
-    i += 1000

### 2014-11-07 article.tex
Integrate online learning example
Integrate on-line learning example

--- 2014-11-06 article.tex
@@
 
+\begin{figure*}[b]
+% To have the float on the bottom, we need this declared on the page
+% before where it is going to be printed
+\lstinputlisting{online_learning_example.py}
+\caption{On-line learning example: the main loop iterates through batches
+of 1000 documents, vectorizing them on the fly and updating the
+classifier. Running the example shows that the error on the test set
+progressively decreases as data is accumulated.\label{fig:online}}
+\end{figure*}
+
 We can test our classifier on two new movie reviews.\footnote{
@@
 the bigger, the less likely collisions will occur.
+%
+A full example doing online learning on a large corpus of text is given
+on figure \ref{fig:online}.
 

### 2014-11-07 online_learning_example.py

@@ -8,7 +8,7 @@ X_test = vectorizer.fit_transform(dataset.data[:1000])
 y_test = dataset.target[:1000]
 
 from sklearn.linear_model import SGDClassifier
-classifier = SGDClassifier()     # SVM classifier that learns online using SGD.
+classifier = SGDClassifier()    # SVM classifier trained online with stochastic gradient descent.
 
 # Artificially make long train data
 train_data = 10 * dataset.data[1000:]
@@ -17,6 +17,6 @@ train_target = 10 * list(dataset.target[1000:])
 for i in range(0, len(train_data), 1000):    # Iterate on mini-batchs of 1000 samples
     y_train = train_target[i:i + 1000]
     X_train = vectorizer.fit_transform(train_data[i:i + 1000])
-    # update estimator with examples in the current mini-batch
+    # update the classifier with documents in the current mini-batch
     classifier.partial_fit(X_train, y_train, classes=range(len(dataset.target_names)))
     print classifier.score(X_test, y_test)

### 2014-11-07 article.tex
Fixing typos and cosmits

--- 2014-11-06 article.tex
@@
 do sentiment polarity classification for movie reviews: given a review,
-it outputs whether that review is positive about the movie it concerns.
+it outputs whether the review is positive about the movie it concerns.
 The code demonstrates the use of pipelines, and also shows how machine

### 2014-11-07 nlp_example.py

@@ -11,9 +11,9 @@ clf = pipeline.Pipeline([('step1', feature_extraction.text.TfidfVectorizer(min_d
 clf.fit(data.data, data.target)
 
 # Demo
-clf.predict_proba(["Worst movie I ever saw"])[0, 0]  # negative probability
+clf.predict_proba(["Worst movie I ever saw"])[0, 0]  # Negative probability
 # 0.99929299972040175
-clf.predict_proba(["Godzilla, eat your heart out!"])[0, 1]  # positive probability
+clf.predict_proba(["Godzilla, eat your heart out!"])[0, 1]  # Positive probability
 # 0.64533559029803611
 print clf.predict(["This movie is the worst I ever saw.",
 "Shawshank Redemption, eat your heart out!"])

### 2014-11-07 online_learning_example.py

@@ -1,22 +1,22 @@
 from sklearn.datasets import fetch_20newsgroups
 from sklearn.feature_extraction.text import HashingVectorizer
 
-vectorizer = HashingVectorizer(decode_error='ignore', n_features=2 ** 16) # A vectorizer
+vectorizer = HashingVectorizer(decode_error='ignore', n_features=2 ** 16)  # A vectorizer
 
 dataset = fetch_20newsgroups(shuffle=True, random_state=1, remove=('headers', 'footers'))
 X_test = vectorizer.fit_transform(dataset.data[:1000])
 y_test = dataset.target[:1000]
 
 from sklearn.linear_model import SGDClassifier
-classifier = SGDClassifier()    # SVM classifier trained online with stochastic gradient descent.
+classifier = SGDClassifier()  # SVM classifier trained online with stochastic gradient descent
 
 # Artificially make long train data
 train_data = 10 * dataset.data[1000:]
 train_target = 10 * list(dataset.target[1000:])
 
-for i in range(0, len(train_data), 1000):    # Iterate on mini-batchs of 1000 samples
+for i in range(0, len(train_data), 1000):  # Iterate on mini-batchs of 1000 samples
     y_train = train_target[i:i + 1000]
     X_train = vectorizer.fit_transform(train_data[i:i + 1000])
-    # update the classifier with documents in the current mini-batch
+    # Update the classifier with documents in the current mini-batch
     classifier.partial_fit(X_train, y_train, classes=range(len(dataset.target_names)))
     print classifier.score(X_test, y_test)

### 2014-11-07 article.tex
online learning example: copyedit, Py3 compat, footnote

--- 2014-11-06 article.tex
@@
 \lstinputlisting{online_learning_example.py}
-\caption{On-line learning example: the main loop iterates through batches
+\caption{On-line learning: the main loop iterates over batches
 of 1000 documents, vectorizing them on the fly and updating the
-classifier. Running the example shows that the error on the test set
-progressively decreases as data is accumulated.\label{fig:online}}
+classifier. Running the example shows the error on the test set
+progressively decreasing as data is accumulated.\label{fig:online}}
 \end{figure*}
@@
 %
-A full example doing online learning on a large corpus of text is given
-in Fig.~\ref{fig:online}.
+A full example doing online learning on the 20 Newsgroups corpus\footnote{\url{http://qwone.com/~jason/20Newsgroups/}}
+is given in Fig.~\ref{fig:online} (a downloader for this corpus comes with scikit-learn).
 

### 2014-11-07 online_learning_example.py

@@ -1,7 +1,7 @@
 from sklearn.datasets import fetch_20newsgroups
 from sklearn.feature_extraction.text import HashingVectorizer
 
-vectorizer = HashingVectorizer(decode_error='ignore', n_features=2 ** 16)  # A vectorizer
+vectorizer = HashingVectorizer(decode_error='ignore', n_features=2 ** 16)
 
 dataset = fetch_20newsgroups(shuffle=True, random_state=1, remove=('headers', 'footers'))
 X_test = vectorizer.fit_transform(dataset.data[:1000])
@@ -10,13 +10,13 @@ y_test = dataset.target[:1000]
 from sklearn.linear_model import SGDClassifier
 classifier = SGDClassifier()  # SVM classifier trained online with stochastic gradient descent
 
-# Artificially make long train data
+# Artificially increase the size of the training set ten-fold
 train_data = 10 * dataset.data[1000:]
 train_target = 10 * list(dataset.target[1000:])
 
-for i in range(0, len(train_data), 1000):  # Iterate on mini-batchs of 1000 samples
+for i in range(0, len(train_data), 1000):  # Iterate over "mini-batches" of 1000 samples each
     y_train = train_target[i:i + 1000]
     X_train = vectorizer.fit_transform(train_data[i:i + 1000])
     # Update the classifier with documents in the current mini-batch
     classifier.partial_fit(X_train, y_train, classes=range(len(dataset.target_names)))
-    print classifier.score(X_test, y_test)
+    print(classifier.score(X_test, y_test))
